### 2022-09-15 src/tex/ms.tex
[showyourwork] overleaf sync

% template.tex, dated April 5 2013
% This is a template file for Annual Reviews 1 column Journals
%
% Compilation using ar-1col.cls' - version 1.0, Aptara Inc.
% (c) 2013 AR
%
% Steps to compile: latex latex latex
%
% For tracking purposes => this is v1.0 - Apr. 2013

\documentclass[letterpaper]{ar-1col}
\usepackage{showyourwork}
\usepackage[letterpaper]{geometry}

\usepackage{natbib}
\usepackage{amsmath}
\usepackage{color}
\usepackage{hyperref}
\hypersetup{hidelinks}

\usepackage{graphbox}
\newcommand{\suz}[1]{\textcolor{magenta}{#1}}
\newcommand{\dan}[1]{\textcolor{green}{#1}}

\setcounter{secnumdepth}{4}
\usepackage{url}

% Metadata Information
\jname{Xxxx. Xxx. Xxx. Xxx.}
\jvol{AA}
\jyear{2022}
\doi{10.1146/((please add article doi))}

% autoref formatting
\def\sectionautorefname{Section}
\let\subsectionautorefname\sectionautorefname
\let\subsubsectionautorefname\sectionautorefname

% macros
\newcommand{\apjl}{Astrophysical Journal Letters}
\newcommand{\aj}{Astronomical Journal}
\newcommand{\apj}{Astrophysical Journal}
\newcommand{\apjs}{Astrophysical Journal Supplement}
\newcommand{\pasp}{Publications of the Astronomical Society of the Pacific}
\newcommand{\jgr}{Journal of Geophysical Research}
\newcommand{\aap}{Astronomy and Astrophysics}
\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
\newcommand{\actaa}{Acta Astronomica}
\newcommand{\nat}{Nature}
\newcommand{\prl}{Physical Review Letters}
\newcommand{\prd}{Physical Review D}

% Symbols
\newcommand{\ydata}{\ensuremath{\boldsymbol{y}}}
\newcommand{\hyperparams}{\ensuremath{\boldsymbol{\phi}}}
\newcommand{\meanparams}{\ensuremath{\boldsymbol{\theta}}}
\newcommand{\dt}{\ensuremath{\tau}}
\newcommand{\amplitude}{\ensuremath{\alpha}}
\newcommand{\lengthscale}{\ensuremath{\lambda}}

\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand{\project}[1]{\textsf{#1}}

% Document starts
\begin{document}

% Page header
\markboth{Aigrain \& Foreman-Mackey}{GPR for Astronomy}

% Title
\title{Gaussian Process regression for astronomical time-series}

%Authors, affiliations address.
\author{Suzanne Aigrain,$^1$ and Daniel Foreman-Mackey$^2$
  \affil{$^1$Department of Physics, University of Oxford, Oxford, UK, OX1 3RH; email: suzanne.aigrain@physics.ox.ac.uk}
  \affil{$^2$Center for Computational Astrophysics, Flatiron Institute, New York, USA, NY 10010}}

%Abstract
\begin{abstract}
  Abstract text, approximately 150 words.
\end{abstract}

%Keywords, etc.
\begin{keywords}
  keywords, separated by comma, no full stop, lowercase
\end{keywords}
\maketitle

%Table of Contents
\tableofcontents

\section{INTRODUCTION}
\label{sec:intro}

Gaussian Processes (GPs) are a powerful class of statistical models, which allow us to define a probability distribution over random functions. Rather than write down an explicit mathematical formula for the function from which some observations are generated, we model the covariance between pairs of samples from the process, using our physical domain knowledge and/or available data to guide our modelling.
While this may seem abstract, GPs have a wide range of applications from modelling of stochastic physical processes, to high dimensional interpolation and smoothing.
\begin{armarginnote}[]
  \entry{GP}{Gaussian Process}
  \entry{GPR}{Gaussian Process Regression}
\end{armarginnote}
In particular, GP Regression (GPR) has become increasingly popular in the astronomical community over the last decade. Part of the reason for this uptake is the growing availability and importance of time-domain datasets in astronomy. These systematically contain non-trivial random or unknown signals, whether astrophysical or instrumental, that need to be modelled. In many cases, these are nuisance signals, which we need to marginalise over in order to detect or measure other signals robustly. Sometimes, we are interested in the stochastic behaviour itself, and want to infer its characteristics or predict its behaviour. GPR offers a compelling solution: statistically principled, naturally Bayesian, and extremely flexible, yet mathematically simple. To new users, however, the lack of an explicit functional form for the model can make GPR can seem a little arcane. Furthermore, the computational cost of the method, which scales cubically with the dataset size, can be an obstacle. These factors initially impeded its dissemination in the astronomical community, but have been largely overcome in recent years thanks to the availability of user-friendly, computationally optimised software packages.

In this review, we aim to give the reader a concise introduction to the basic theoretical framework of GPR, and to illustrate the range, strengths and limitations of the method by discussing some of its applications to time-domain datasets in astronomy so far. We made no attempt to provide a complete theoretical description or to discuss the relevant literature exhaustively: there are excellent textbooks which already do the former, while the latter would have been impossible in the space available. Instead our goal was to provide an accessible introduction that might give interested readers starting points for further reading, and help them determine whether GPR might be useful for their own datasets. With this in mind, we took pains to require minimal prior knowledge, to provide practical advice regarding the modelling choices one needs to make when using GPR, and to review relevant open-source software.

% In this review, we start with a broad and pedagogical introduction to GP methodology in theory and practice.
% We then work through many examples of how GPs have been used in the astronomical literature.
% Discuss

As a technical note, this manuscript was prepared using the \project{showyourwork} package\footnote{\url{https://show-your.work}} and the source code used to generate each figure is available in a public \project{GitHub} repository\footnote{\url{https://github.com/dfm/araa-gps}}.
To see the specific version of the \project{Jupyter} notebook, that was executed to generate each figure, click on the icon next to the figure caption.

\subsection{Brief history}

An early use of GPR, was for spatial interpolation in geophysics \citep{kriging}, and GPR has since been adopted or re-invented in a wide range of other application domains. GPs were used in simulations in a wide range of astronomical sub-fields \citep[see e.g.][]{1980asfr.symp..159B,1988JGR....9311569C,1997ApJ...483L...1P}, but early mentions of GPs for modelling astronomical datasets \citep[see e.g.][]{1976MitAG..38..192D,1978A&A....70..777V,1991MGeo...16..313J} received limited attention.

Perhaps the earliest use of GPR in the refereed astronomical literature that will be familiar to a modern reader was published by \citet{prh92a} in the context of quasar variability, and for a long time this remained its main application domain in astronomy. GPR then gradually appeared in other areas, starting with photometric redshift estimation \citep{2006ApJ...647..102W}, then exoplanet transit observations \citep{2009ApJ...704...51C,2012MNRAS.419.2683G} and radial velocity planet searches \citep{2012MNRAS.419.3147A,2014MNRAS.443.2517H}. Nonetheless, GPR remained relatively niche and few astronomers had heard of it until a few years ago. To illustrate this point, we searched on the NASA Astrophysics Data System (ADS) for articles published in refereeed astronomy and astrophysics journals with the words ``Gaussian Process'' in full text of the article \autoref{fig:literature}.
After an increase in popularity throughout the 1990s, the use of GPs in astrophysics remained fairly constant around $\sim20$ publications per year until 2010.
Since 2010, the popularity of GPs has grown significantly, and in 2021, more than 500 refereed papers referencing GPs were published in the astrophysics literature.

A number of important factors have contributed to recent the democratisation of GPR across a wide range of scientific disciplines, including the publication of a dedicated textbook \citep{gpml}, as well as the availability of cheap computing power and user-friendly, open-source GPR software. In astronomy, specifically, an additional factor has been at play: the rise of time domain surveys. Correlated noise in time-domain observations is a direct and unavoidable consequence of causation, and hence ubiquitous. Adequately modelling this correlated noise is vital when searching for faint signals, for example from exoplanets. Astrophysical sources, from accretion disks on all scales to magnetically active stars or cloudy brown darfs, also display complex, intrinsically or apparently stochastic behaviour, for which adequate modelling strategies are required. GPR is a natural choice to tackle these challenges. Rather than attempting to cover all applications of GPs to all of astrophysics, which would not be feasible in the space available, we have therefore opted to focus on its application to time-domain datasets in astronomy.

\begin{figure}[ht]
  \centering
  \script{literature.ipynb}
  \includegraphics[width=0.5\linewidth]{figures/literature.pdf}
  \caption{The number of referred publications in the astronomy and astrophysics literature that include the text ``Gaussian process'' as indexed by the NASA Astrophysics Data System (ADS). \suz{Add paper network?}}
  \label{fig:literature}
\end{figure}

\subsection{Motivating examples}
\label{sec:sim_examples}

By way of motivation, before delving any further into the methodology, we have produced two illustrative examples that highlight some typical use cases for GPR in astrophysics.
These examples are not meant to comprehensively summarize the space of use cases.
Similarly, the goal of this section is not to formally compare the performance of GPR to other methods with the same goals.
Instead, these examples are meant to identify some common, but qualitatively different applications.
In both examples, we use simulated datasets since it is useful to know the ``ground truth'' to validate performance.

The first example is a re-implementation of one of the earliest uses of GPR in time-domain astronomy \citep{prh92a}, using modern language and techniques.
In this example, we measure the time delay of a lensed quasar, using a GP as a flexible, non-parametric model for the latent (un-observed) variability of the unlensed quasar system.

The second example demonstrates the use of GPR to account for stellar variability in the light curve of a transiting exoplanet, when inferring its parameters.
In this case, the parameters of interest are the parameters of the mean model, and the GP is a nuisance model, and our goal is to propagate uncertainty introduced by the stellar variability to our constraints on the physical parameters of interest.

These examples---and all the examples throughout this review---are implemented using \textsf{tinygp}\footnote{\url{https://tinygp.readthedocs.io}} \citep{tinygp}, a \project{Python} library for GPR built on top of the \project{JAX}\footnote{\url{https://jax.readthedocs.io}} library for numerical computing \citep{jax}.
Here and throughout, the probabilistic models are implemented using the \project{NumPyro}\footnote{\url{https://num.pyro.ai}} library \citep{numpyro} and the Markov chain Monte Carlo (MCMC) inference is performed using the No U-Turn Sampling (NUTS) algorithm \citep{Hoffman:2014}.

\subsubsection{Example 1: The time delay of a gravitational lensed quasar}
\label{sec:quasar}

In this example, we re-visit the method developed by \citet{prh92a} to measure the time delay of the gravitationally lensed quasar 0957+561, one of the earliest applications of GPR for time domain astronomy.
The underlying model here is that the unobserved latent variability of the source is modelled using a GP, in this case we use a Mat\'ern-3/2 covariance function as discussed and defined in \autoref{sec:kernels}.
The images sample this time series at lagged times and with different mean magnitudes and variability amplitudes.

Under this assumed model, we simulate a pair of light curves with the same cadence and uncertainties as the dataset from \citet{1989A&A...215....1V} that was analyzed by \citet{prh92a}.
In this simulation, the parameters of the covariance model, the mean magnitudes, and the time delay are all set to known values, designed to produce qualitatively similar features to the dataset from \citet{1989A&A...215....1V}.
The simulated light curves are plotted in the left panel of \autoref{fig:quasar}.

Using these simulated data, we fit a GP model using MCMC, varying the time delay, the mean magnitude of each image, the variability amplitude of each image, and the timescale of the covariance.
The results of this inference are shown in \autoref{fig:quasar}.
In the left panel we show the simulated data with the median of posterior time delay applied to image A, and an arbitrary magnitude offset applied to image B for plotting purposes.
Over-plotted on these data are 12 posterior samples of the GP model predictions for the noise-free photometry for each image.
This figure captures how a GP can be used to flexibly capture a stochastically variable process under certain smoothness constraints, and how the uncertainties on the interpolated and extrapolated predictions increase away from the observed data.

The right panel of \autoref{fig:quasar} shows the posterior constraints for two of the key parameters of the model: the time delay, and the mean magnitude difference between images.
Since these are simulated data, we know the true values of these parameters, and these true values are over-plotted on \autoref{fig:quasar}b, demonstrating that our method reliably recovers the expected result.

\begin{figure}[ht]
  \centering
  \script{quasar.ipynb}
  \begin{minipage}[t]{0.5\linewidth}
    \includegraphics[width=\linewidth,vshift=1cm]{figures/quasar.pdf}
  \end{minipage} \hfill
  \begin{minipage}[t]{0.44\linewidth}
    \includegraphics[width=\linewidth]{figures/quasar_posteriors.pdf}
  \end{minipage}
  \caption{The results of fitting the simulated light curves of two images of a lensed quasar with a time delay between images.
  \emph{Left:} The simulated light curves for each image with an arbitrary magnitude offset applied to image B, and the median of posterior time delay applied to image A, bringing the light curves into a common frame.
  The data are plotted as points with error bars, and over-plotted on these data are posterior predictive samples of each images' predicted variability.
  \emph{Right:} The posterior constraints on the time delay and mean magnitude offset between the two images, obtained using Markov chain Monte Carlo (MCMC) to fit the simulated data shown in the left panel of this figure.
  The true values of these parameters that were used to simulate the data are over-plotted as blue lines.}
  \label{fig:quasar}
\end{figure}

\subsubsection{Example 2: Fitting an exoplanet transit with stellar variability}
\label{sec:transit}

In this second example, we demonstrate another common application of GPR in astrophysics, as a flexible model for nuisances and correlated noise where we want to correctly capture uncertainty introduced by this noise model into our constraints on the parameters of interest.
To this end, we simulate the light curve of a transiting exoplanet and aim to infer the physical parameters of the system, taking correlated noise into account.
The correlated noise in transit light curves is typically caused by the variability of the host star and by instrumental effects like focus or pointing changes.
It has been demonstrated that neglecting to account for these variations can cause significant errors when inferring the properties of the planet \citep{2006MNRAS.373..231P,2007A&A...472L..13G}.
This correlated noise can often be well-modeled by a GP, and the use of GPs for transit modeling has been a fruitful area of research in the astrophysics literature (see Section~\ref{sec:transit_fit} for a more detailed discussion).

For this example, the transit was simulated with known physical properties such as the planet-to-star radius ratio, the impact parameter, and quadratic limb darkening parameters \citep{2020AJ....159..123A}.
We then add correlated and white noise by sampling from a GP model with a Mat\'ern-3/2 covariance function as defined in \autoref{sec:kernels}, with a known amplitude, timescale, and white noise amplitude.
These simulated data are shown in the left panel of \autoref{fig:transit}.

We then used the same framework to model the simulated dataset, first ignoring the correlated noise and then accounting for it explicitly.
For the transit model, we fit for the planet-to-star radius ratio $R_{\rm p}/R_\star$, the mid-transit time $T_0$, the out-of-transit flux $f_0$, and the limb darkening parameters $u_1,\,u_2$.
When also fitting for correlated noise, we also marginalise over the amplitude $\amplitude$ and timescale $\lengthscale$ of the GP model.
The posterior constraints on the time of transit and the planet-to-star ratio are shown in the right panel of \autoref{fig:transit}, with the true values of these parameters over-plotted as a black line.
When the correlated noise is neglected (plotted in orange in \autoref{fig:transit}), the inferred parameters are significantly inconsistent with the truth.
Taking the correlated noise into account (blue in \autoref{fig:transit}) increases the uncertainties on the inferred parameters, but also shifts the results to recover the true parameters.

\begin{figure}[ht]
  \centering
  \script{transit.ipynb}
  \begin{minipage}[t]{0.5\linewidth}
    \includegraphics[width=\linewidth,vshift=1cm]{figures/transit.pdf}
  \end{minipage} \hfill
  \begin{minipage}[t]{0.44\linewidth}
    \includegraphics[width=\linewidth]{figures/transit_posteriors.pdf}
  \end{minipage}
  \caption{The results of fitting the simulated light curve of a transiting exoplanet.
  \emph{Left:} The simulated dataset plotted as points with error bars.
  The inferred model when correlated noise is (blue) or isn't (orange) taken into account are over-plotted on the data.
  \emph{Right:} The posterior constraints on the time of transit $T_0$ and the planet-to-star radius ratio $R_\mathrm{p}/R_\star$.
  The colors in this figure match those in the left panel, and the true values of these parameters are indicated with black lines.
  The results when neglecting correlated noise (orange) are significantly inconsistent with the true value, but when the correlated noise is modeled using a GP (blue), the correct parameters are recovered, albeit with larger uncertainty.}
  \label{fig:transit}
\end{figure}

\subsection{Overview of this review}

The remainder of this review is structured as follows. Section~\ref{sec:basics} provides a brief, accessible introduction to the basic theory of GPR, and Section~\ref{sec:choices} discusses the key modelling choices one needs to make when using a GP to model data, with practical advice for how to make these choices. Section~\ref{sec:uses} gives an overview of applications of GPR to time-domain astronomical datasets to date, from stars and exoplanets to active galactic nuclei, pulsars and gravitational waves. In Section~\ref{sec:challenges} we discuss some of the challenges of using GPR in practice, and how they may be overcome, while Section~\ref{sec:open} reviews the ecosystem of open-source GPR software available today. Finally, we summarise the main take-away points and outline future directions in Section~\ref{sec:concl}.

% Heading 1
\section{BASICS OF GAUSSIAN PROCESS REGRESSION}
\label{sec:basics}

So far we have discussed how GPR has become widely used in astronomy and touched on some of their pros and cons, but we haven't explicitly defined what a GP is or explained how GPR works. This section gives a brief introduction to the theory of GPR, written in a way that we hope will be intelligible to most astronomers. A much more in-depth treatment is provided by \citet{gpml}, which to this day remains the main reference textbook on GPs.

\subsection{Formal definition}
\label{sec:def}

A GP is a type of \textit{stochastic process} based on the Gaussian probability distribution. A probability distribution describes a random variable with a finite number of dimensions. A stochastic process extends this concept to an infinite number of dimensions, allowing us to define a probability distribution over functions. Just what do we mean by ``extending to an infinite number of dimensions"? Well, this can be a little problematic mathematically, but we needn't worry about it, because in practice we only ever deal with finite samples from the stochastic process.

The formal definition of a GP is that the joint probability distribution over any finite sample $\mathbf{y} = \{y_i\}_{i=1,\ldots,N}$ from the GP is a multi-variate Gaussian:
\begin{equation}
\label{eq:gp_def}
    p(\mathbf{y}) = \mathcal{N}(\mathbf{m}, \mathbf{K}),
\end{equation}
where $\mathbf{m}$ is the \emph{mean vector} and $\mathbf{K}$ the \emph{covariance matrix}.

The elements of the mean vector and covariance matrix are given by the \textit{mean function} $m$ and the \textit{covariance function} $k$, respectively:
\begin{equation}
    m_{i} = m(\boldsymbol{x}_i, \meanparams),
\end{equation}
\begin{equation}
    K_{ij} = k(\boldsymbol{x}_i,\boldsymbol{x}_j, \hyperparams).
\end{equation}
where $\boldsymbol{x}_i$ is the set of inputs (independent variables) corresponding to the $i^{\rm th}$ sample. For time-series data, the inputs usually include, but aren't necessarily restricted to, the time $t_i$. The covariance function, also known as the \emph{kernel function}, is the fundamental ingredient of a GP model, and considerable care must be taken to select it adequately (or to test different possibilities). Sometimes the mean function is assumed to be constant, or even zero, everywhere; this is often done in the wider GPR literature to keep derivations uncluttered. However, for many astrophysical applications where a GP is used to model a nuisance signal, the mean function contains the signal of interest and is important.

\begin{figure}[ht]
  \centering
  \script{GPR_PGM.ipynb}
  \includegraphics[width=0.8\linewidth]{figures/GPR_PGM.pdf}
  \caption{Probabilistic Graphical Model (chain graph) for GP regression (adapted from Figure 2.3 of \citealt{gpml}) given observations $\mathbf{y}$ at inputs $\mathbf{t}$, and a prediction at test input $t_\star$. Dots represent fixed variables (inputs), grey circles represent observed variables white circles represent unknown (latent) variables. The thick thick horizontal bar indicates a set of fully connected nodes (meaning that every $f$ depends on every $t$). On the other hand, each observed output $y_i$ depends only on the corresponding $f_i$ and $\sigma_i$, and is conditionally independent of the other variables). In this example, the measurement uncertainties $\boldsymbol{\sigma}$ and the GP hyper-parameters (not shown) are treated as known (fixed).}
  \label{fig:GPR_PGM}
\end{figure}

The parameters $\meanparams$ and $\hyperparams$ of the mean and covariance function are known as the \textit{hyper-parameters} of the GP. Strictly speaking, the \textit{parameters} of the GP are the (infinitely many) unknown functions that share the specified mean vector and covariance matrix and could have given rise to the observations. However, these parameters are always marginalised over: we never explicitly deal with the individual functions, except when drawing samples for illustrative purposes (as we did in Figures~\ref{fig:quasar} and \ref{fig:transit}). GPs are therefore a type of \textit{Hierarchichal Bayesian Model} (HBM). To illustrate this, we show a Probabilistic Graphical Model (PGM) of GPR in Figure~\ref{fig:GPR_PGM}. In this type of graph, shaded circles represent random variables that are observed, while un-shaded circles represent latent, or unobserved, random variables. We never observe the unknown function that gave rise to the data directly, but infer a probability distribution for it from our noisy observations.

Although is is possible to construct and use stochastic process models based on other distributions, GPs are by far the most popular, for two main reasons. The first is Central Limit theorem: it implies that the assumption of Gaussianity is often at least approximately correct. The second is that Gaussian distributions obey simple mathematical identities for marginalisation and conditioning, that enable inference with GPs (the process of marginalising over the individual functions) to be performed \textit{analytically}, with very simple linear algebra. It is this analytic marginalisation property that sets GPR apart from other forms of HBM.

\subsection{From least-squares to GPR}
\label{sec:from_lsq}

The formal definition given in Equation~\ref{eq:gp_def} does not necessarily provide an intuitive understanding of how GPs work. Most readers of this review will, however, be more familiar with least-squares regression. In this section we will show that GPR can be thought of as a generalisation of least-squares regression, allowing for correlated noise (or signals) in the data. Conversely, least-squares regression, as traditionally presented, is a special case of GPR, where the covariance matrix is assumed to be purely diagonal, and the variances associated with each observation are known \textit{a priori}.

\subsubsection{Revisiting least squares}\label{sec:lsq}

Once again, we consider $N$ observations of a variable $\mathbf{y} = \{ y_i\}_{i=1, \ldots, N}$, taken at times $\mathbf{t} = \{ t_i\}$, with associated measurement uncertainties $\boldsymbol{\sigma}=\{\sigma_i\}$. We wish to compare these to a model function $m(t,\meanparams)$ controlled by parameters $\meanparams=\{\theta_j\}_{j=1,\ldots,M}$. In most astrophysical applications, we are interested in estimating some or all of those parameters. In least-squares regression, we minimize the quantity
\begin{equation}
  \chi^2 \equiv\sum_{i=1}^N (y_i-m_i)^2/\sigma^2_i,
\end{equation}
where $m_i \equiv m(t_i,\meanparams)$, with respect to $\meanparams$. Where does this come from?

Let us assume that the observations are given by
\begin{equation}
  y_i=m(t_i,\meanparams)+\epsilon_i,
\end{equation}
where $\epsilon_i$ is the measurement error, or noise, on the $i^{\rm th}$ observation. Furthermore, let us assume that $\epsilon_i$ is drawn from a Gaussian distribution with mean $0$ and variance $\sigma_i^2$:
\begin{equation}
  p(\epsilon_i)=\mathcal{N}(0,\sigma_i^2) \equiv \frac{1}{\sqrt{2\pi} \sigma_i} \exp\left(-\frac{\epsilon_i^2}{2\sigma_i^2} \right),
\end{equation}
then the \emph{likelihood} for the $i^{\rm th}$ observation is simply
\begin{equation}
  \mathcal{L}_i (\meanparams) \equiv p(y_i|\meanparams)=\mathcal{N}(m_i,\sigma_i^2) =\frac{1}{\sqrt{2\pi} \sigma_i} \exp\left[-\frac{(y_i-m_i)^2}{2\sigma_i^2} \right].
\end{equation}
Furthermore, we also assume that the noise is uncorrelated, or white, meaning that the $\epsilon_i$'s are drawn independently from each other from their respective distributions. Then, the likelihood for the whole dataset $\mathbf{y}$ is merely the product of the likelihoods for the individual observations:
\begin{equation}
  \mathcal{L} (\meanparams) \equiv p(\mathbf{y}|\meanparams) = \prod_{i=1}^N \mathcal{L}_i = \prod_{i=1}^N \left\{ \frac{1}{\sqrt{2\pi} \sigma_i}
  \exp\left[-\frac{(y_i-m_i)^2}{2\sigma_i^2} \right] \right\},
\end{equation}
where $\mathbf{m}=\{m_i\}_{i=1,\ldots,N}$. From this one can readily see that $\ln \mathcal{L} = \mathrm{constant} - 0.5 \chi^2$, where the constant depends only on the $\sigma$'s. Thus, if the $\sigma$'s are known, maximizing $\mathcal{L}$ is equivalent to minimizing $\chi^2$. In other words, least-squares regression yields the Maximum Likelihood Estimate (MLE) of the parameters under the assumption of white, Gaussian noise with known variance.

\subsubsection{Link to Gaussian Process Regression}\label{sec:gp-lsq-link}
% \dan{I think we should explicitly re-write this as $\mathcal{L}(\meanparams,\,\hyperparams)$ and indicate that $K$ and $m$ depend on these parameters. We should really highlight that this is a drop-in replacement for the likelihood defined in the previous section.} \suz{Is this ok?}
Let us now re-write the likelihood in matrix form:
\begin{equation}
\label{eq:gp_like}
  \mathcal{L}(\meanparams,\hyperparams) = 
  \frac{1}{\sqrt{2\,\pi\,|\mathbf{K}|}} \exp \left(-(\mathbf{y}-\mathbf{m})^{\mathrm{T}} \mathbf{K}^{-1}
  (\mathbf{y}-\mathbf{m}) \right),
\end{equation}
where, as before, the mean vector $\mathbf{m}$ has elements $m_i=m(t_i,\meanparams)$ and $\mathbf{K}$ is a purely diagonal ($N$, $N$) matrix with elements $K_{ij} = \delta_{ij} \sigma_i^2$ ($\delta_{ij}$ being the discrete Kronecker delta function). This is, of course, the \emph{covariance matrix} of the model. The notation $|\mathbf{A}|$ refers to the determinant of the matrix $\mathbf{A}$, and $\mathbf{A}^{-1}$ to its matrix inverse.

Now, instead of assuming that the covariance matrix takes this very specific form, let us allow a more flexible covariance model:
\begin{equation}
  K_{ij} = k(t_i,t_j,\hyperparams) + \delta_{ij} \sigma_i^2,
\end{equation}
where $k$ is a \emph{covariance function}, or \emph{kernel function}, controlled by parameters $\hyperparams$. The result is a GP, and its likelihood is still given by Equation~\ref{eq:gp_like}. Depending on the choice of kernel function and parameters, the covariance matrix can now have non-zero off-diagonal elements, allowing us to explicitly model correlated noise or stochastic signals in the data.
Note that the likelihood depends not only on the argument of the exponential, but also on the determinant of the covariance matrix, which therefore needs to be evaluated explicitly. This term acts as a built-in Occam's razor, automatically penalising more complex models.
%\dan{We should probably define the determinant notation here? I don't think we really clearly define our matrix notation anywhere else.}

The kernel function $k$ encodes our beliefs about the stochastic, or random, element of the model, in just the same way as the mean function $m$ encodes our beliefs about the deterministic component of the model. For example, in many circumstances, we would expect that two observations taken close together in time should be more strongly correlated than observations taken further apart, and we would use a decreasing function of the time interval $\dt=|t_i-t_j|$ to represent that. One of the most widely used covariance functions is the squared exponential
\begin{equation}
\label{eq:kse}
  k(\dt;\,\hyperparams) = \amplitude^2\,\exp\left(-\frac{\dt^2}{2\,\lengthscale^2}\right)
\end{equation}
This gives rise to smooth random functions with variance $\amplitude$ and characteristic length-scale $\lambda$. We present other commonly used kernel functions and discuss how to select one in Section~\ref{sec:kernels}.

\subsection{Inference with a GP}\label{sec:gp-inf}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{figures/GPR_workflow.pdf}
  \caption{Schematic representation of a typical GPR workflow. Given a dataset (blue box) and some modelling choices (grey boxes, Section~\ref{sec:choices}), the mathematical framework presented in Section~\ref{sec:basics} can be used (green boxes) to evaluate the likelihood and posterior distribution over the HPs which can be optimized or sampled. We can also condition the GP on the data to predict observations at locations where we do not yet have observations. Several of the plots in this figure are taken from the 
  tutorial on hyper-parameter optimization included in the {\sc george} package documentation (\url{https://george.readthedocs.io}), and show the composite GP model proposed by \citet{gpml} for the Mauna Loa CO$_2$ dataset of \citet{MaunaLoaCO2}.}  
  \label{fig:GPR_workflow}
\end{figure}

Now that we know how to evaluate the likelihood (Equation~\ref{eq:gp_like}), we are ready to perform \textit{inference}, that is to use observations to update our prior beliefs about the system we are observing. The overall Bayesian inference workflow for GPR regression is illustrated schematically in Figure~\ref{fig:GPR_workflow}. Given a dataset (blue, left panel), our first task is typically to select a kernel function, along with an educated initial guess for its hyper-parameters ($2^{\rm nd}$ column, top panel). These choices are discussed in more detail in Section~\ref{sec:choices}, and should be guided by a combination of domain knowledge (what we know about the process that generated the data), practical considerations (e.g.\ ease of implementation and computational cost) and diagnostics based on the data themselves (e.g.\ examining the auto-correlation function and/or the power spectral density of the data). Given a kernel function and hyper-parameters, we can draw samples from the GP prior ($3^{\rm rd}$ column, top panel), which can serve as sanity check for the choice of kernel function (in the sense that, measurement errors aside, samples from the GP prior should ideally look qualitatively similar to the observations).

We are now in a position to compute the GP covariance matrix for the data ($2^{\rm nd}$ column, $2^{\rm nd}$ panel), and to evaluate the likelihood of the model ($2^{\rm nd}$ column, $3^{\rm rd}$ panel), which we can then optimize with respect to the hyper-parameters ($3^{\rm rd}$ column, $2^{\rm nd}$ panel) using standard optimization methods; this is called \emph{training} the GP. Often, however, we would incorporate priors on the hyper-parameters ($2^{\rm nd}$ column, bottom panel), which we select once again based on domain knowledge and practical considerations. We can then evaluate the \emph{posterior} distribution ($3^{\rm rd}$ column, $3^{\rm rd}$ panel):
\begin{equation}
  p(\meanparams,\hyperparams|\mathbf{y})=\frac{p(\mathbf{y}|\meanparams,\hyperparams)p(\meanparams,\hyperparams)}{p(\mathbf{y})},
\end{equation}
where the denominator $p(\mathbf{y})=\int p(\mathbf{y}|\meanparams,\hyperparams)p(\meanparams,\hyperparams) \mathrm{d}\meanparams \mathrm{d}\hyperparams$ is the marginal likelihood, or model evidence, and is often omitted from calculations since it is constant, but becomes relevant for model comparison. Once again, we can optimize this posterior with respect to the hyper-parameters ($3^{\rm rd}$ column, $2^{\rm nd}$ panel). For a given set of hyper-parameters (for example, the set which maximises either the likelihood or the posterior), the GP prior can be conditioned on the observations to make predictions ($4^{\rm th}$ column, top panel); this is called \emph{conditioning} the GP, and is discussed in Section~\ref{sec:pred}.

However, in astronomical applications we often need to estimate the full posterior distribution over the hyper-parameters. 
Generally we cannot do this analytically, so we resort to sampling methods such as Markov Chain Monte Carlo (MCMC) or nested sampling ($3^{\rm rd}$ column, bottom panel). The samples can be used to estimate posterior distributions for individual hyper-parameters of interest ($4^{\rm th}$ column, $2^{\rm nd}$ panel). For example, we might be using a GP to take into account correlated noise when fitting an otherwise deterministic model to some data, but we are not interested in the noise \emph{per se}, only on the impact it has on the parameters of the mean function, $\meanparams$. We would then \emph{marginalize} (integrate over, or project) the posterior samples over the \emph{nuisance parameters} $\hyperparams$. Some samplers also enable us to compute the model evidence, which can be used to compare different kernel functions (see Section~\ref{sec:model_selection} for a discussion). Finally,  posterior samples can also be used in a Monte Carlo fashion to compute predictions that incorporate the uncertainties on the HPs ($4^{\rm th}$ column, bottom panel).

\subsection{Making predictions}
\label{sec:pred}

An important use case for GPR is as a statistically principled interpolation method. We wish to ``learn'' an unknown function that gave rise to some data, in order to make predictions for some new set of inputs. This process is illustrated schematically in Figure~\ref{fig:GPR_PGM}, where such a test input and the corresponding function value are highlighted in red. Importantly, a GP model provides not a point estimate but a full probability distribution for the function at any desired location(s) in the input domain. This allows for robust uncertainty propagation (though there are some important caveats we will touch upon in Section~\ref{sec:caution}), and can also motivate strategies for active sampling (deciding when or where to make new observations).

\subsubsection{The predictive equations}

Given some existing observations $\mathbf{y}$, taken at times $\mathbf{t}$, how do we make predictions at some new set of times $\mathbf{t}_\star$? We are after $p(\mathbf{y}_\star|\mathbf{y})$, the \textbf{conditional} probability distribution for $\mathbf{y}_\star$ given $\mathbf{y}$. In GPR this is also known as the \emph{predictive distribution}, because it is often used to extrapolate a time-series dataset forwards. The `magic' of GPR is that the predictive distribution is also Gaussian, and its mean and covariance are given by simple analytic relations:
\begin{equation}
  \label{eq:pred}
  \mathbf{f}_\star = \mathbf{m}_\star + \mathbf{K}_\star^{\mathrm{T}} \, \mathbf{K}^{-1} \, (\mathbf{y} - \mathbf{m}) ~\mathrm{~~~and~~~}
  \mathbf{C}_\star =   \mathbf{K}_{\star\star} - \mathbf{K}_\star^{\mathrm{T}} \, \mathbf{K}^{-1} \, \mathbf{K}_\star,
\end{equation}
where we have introduced the notation $\mathbf{m}_\star \equiv m(\mathbf{t}_\star,\meanparams)$, $\mathbf{K}_\star \equiv k(\mathbf{t},\mathbf{t}_\star, \hyperparams)$ and
$\mathbf{K}_{\star\star} \equiv k(\mathbf{t}_\star,\mathbf{t}_\star,\hyperparams)$. Note that, as real observations are always noisy, $\mathbf{K}$ generally includes a white noise term ($\delta_{ij} \sigma^2_i$), but $\mathbf{K}_\star$ does not. Depending on whether one wishes the predictive variance (the diagonal elements of $\mathbf{C}_\star$)  to account for measurement uncertainties or not, one can include the white noise term in $\mathbf{K}_{\star\star}$ or leave it out.

\subsubsection{Properties of the  predictive distribution}
A closer look at Equation~\ref{eq:pred} reveals some important properties.
First, a GP is a \textit{linear predictor}. The predictive mean for a specific time $t_\star$ can be written as a linear combination of the observations:
$f_\star = \mathbf{w}^{\mathrm{T}} \, \mathbf{y}$, where $\mathbf{w} = \mathbf{K}^{-1} \, \mathbf{k}_\star$ and $\mathbf{k}_\star=k(\mathbf{t},t_\star,\hyperparams)$. It can also be written as
a linear combination of covariance functions centred on the training points:
$f_\star = \mathbf{\alpha}^\mathrm{T} \,  \mathbf{k}_\star$ where $\mathbf{\alpha} = \mathbf{K}^{-1} \, \mathbf{y}$.
These linearity properties are very important to understand the behaviour of GPs. They shed light on the relationship between GPR and standard linear models with very large numbers of free parameters \citep[see e.g.][for a more detailed discussion]{2021PASP..133i3001H},
and lead to a number of powerful extensions that are beyond  scope of this review.

Second, the predictive covariance is independent of the data. $\mathbf{C}_\star$ depends only on the \emph{locations} $\mathbf{t}$ of the observations, not on their values $\mathbf{y}$. This has important consequences for observation planning: if we know the covariance function $k$ and its parameters, we can decide when to observations to optimise our predictions at a given (set of) time(s). However, in practice we rarely know the hyper-parameters \textit{a priori}, and the predictive posterior distribution marginalised over the hyper-parameters \textit{does} depend on the observations.

Finally, as $\mathbf{K}$ is positive semi-definite, so is $\mathbf{K}^{-1}$. Therefore, $\mathbf{k}_\star^{\mathrm{T}} \, \mathbf{K}^{-1} \, \mathbf{k}_\star \ge 0,$ and $\mathrm{Var}(\mathbf{f}_\star) \leq k(t_\star,t_\star)$. This is as we would expect: obtaining more data should only ever improve the accuracy of our predictions.

\subsubsection{Cautionary notes}
\label{sec:caution}
It is important to note that the behaviour of $\mathbf{f}_\star$ is \emph{not} the same as that of individual samples from the predictive distribution. Typically,  $\mathbf{f}_\star$ tends to be smoother than individual samples. This should be borne in mind when displaying GPR results or using them in subsequent analysis.

It is also important to note that the predictive variance accounts for the imperfect ability of the specific model under consideration to explain the data, but not for the choice of model (i.e. the choice of mean and kernel functions and their parameters). We discuss how to choose a kernel function and fit for its parameters in the next section.

\section{GAUSSIAN PROCESS MODELING DECISIONS}
\label{sec:choices}

In the previous section, we presented an overview of GP methods, and the key mathematical details. In this section, we will dive deeper into some of the practical decisions that arise when using GPs.
The two core elements of a GP model are the mean function $m(t;\,\meanparams)$, and the ``kernel'' or ``covariance'' function $k(t_i,\,t_j;\,\hyperparams)$. In contrast to the way GP models are typically presented in the machine learning literature, in astrophysics they will often---but not always---include non-trivial mean functions.
For example, in the case study discussed in \autoref{sec:transit}, the mean function $m(t;\,\meanparams)$ is a physical transit model that is a function of the orbital parameters of the system, and includes a realistic limb-darkening model \citep{2002ApJ...580L.171M}.
However, in this review we won't discuss the mean function in detail, focusing instead primarily on the kernel function, since that is unique to GP modeling.
All this being said, in our experience, new users of GP models will often focus and worry more than necessary about the choice of kernel function for their problem.
As with any probabilistic modeling problem, there are several well-defined workflows for motivating, selecting, and validating your choice of kernel function.
In this section we walk through this process in detail.

\subsection{Gaussian Process Covariance Functions}
\label{sec:kernels}

\begin{textbox}[htb]
  \section{Kernel function}
  The ``kernel'' or ``covariance'' function, typeset here as $k(t_i,\,t_j;\,\hyperparams)$ is a parametric description of the covariance between two data points $t_i$ and $t_j$. This function takes a pair of inputs (here $t_i$ and $t_j$) as arguments, and is controlled by a set of parameters $\hyperparams$, typically referred to as ``hyper-parameters''.

  \vspace{1.5em}

  \section{Hyper-parameters}
  In a GP model, the kernel function is typically parameterized by a set of parameters that we will label as $\hyperparams$ and refer to as ``hyper-parameters''.
  Formally, the ``parameters'' of a GP are the (unknown) ``true'' values of the process at the observed times, but the magic of Gaussians means that these parameters can be marginalised over in closed form, leaving the hyper-parameters as the parameters of interest.
\end{textbox}

The kernel function can be any positive scakar function that gives rise to a positive semi-definite covariance matrix over the input domain, but some will be more useful than others.
Given this large decision space, you may be wondering how to choose the right kernel function for your specific problem.
If you're very lucky, you may be able to motivate your model using physics.
This can be approached from two directions.
On one hand, some commonly used kernel functions have a specific physical interpretation---for example, the solution to a stochastic ordinary differential equation---that would naturally motivate their use in certain circumstances.
On the other hand, if your physical model is specified by its \emph{power spectrum}, this can recast as a GP model with a specific kernel function.
% \suz{Say something about the number of times a function is differentiable? That can be a useful consideration. But perhaps already mentioned later.}

Even if you don't have a formal physics-based justification for your model, you may be able to identify the key scales of your problem and design a kernel function that captures these features.
The usual approach to this problem is to take sums and products of commonly used kernel functions to select a set of  models that have the needed covariance structure, and then combine or select between these choices.
For example, if there is expected to only be a single non-periodic timescale in the problem of interest, you could list all the two-parameter non-periodic kernel functions and use a numerical model selection technique as described below.

\subsubsection{Standard kernel functions}\label{sec:standard-kernels}

Some popular kernel functions are listed in \autoref{tab:kernels}, and some other choices are discussed in Chapter 4 of \citet{gpml}. We already met the 
the squared exponential kernel function, defined in Equation~\ref{eq:kse}. Its  hyperparameters are $\hyperparams = \{\amplitude,\,\lengthscale\}$, where $\amplitude$ controls the amplitude (output scale) and $\lengthscale$ the length scale (input scale) of the functions. Due to its simplicity, this is the most widely-used kernel for modelling smooth functions of unknown shape.

Another example that is commonly used in the astrophysics literature \citep{2012MNRAS.419.3147A,2014MNRAS.443.2517H} is the following ``quasi-periodic'' kernel function
\begin{equation}
\label{eq:kqp}
  k(\dt;\,\hyperparams) = \amplitude^2\,\exp\left(-\frac{\dt^2}{2\,{\lengthscale_1}^2} -\Gamma\,\sin^2\left[\frac{\pi\,\dt}{\lengthscale_2}\right] \right)
\end{equation}
with $\hyperparams = \{\amplitude,\,\lengthscale_1,\,\lengthscale_2,\,\gamma\}$, which has a period of $\lengthscale_2$, and a decoherence timescale of $\lengthscale_1$. The hyper-parameter $\Gamma$ controls the extent to which the periodic component of the signal resembles a simple sinusoid or is more complex.

It's worth noting that, in any practical application, these kernel functions are not generally used on their own.
Instead, more expressive models are designed by combining these models as discussed in \autoref{sec:kernel-ops}.

When selecting a kernel function, it can be useful to generate samples from this implied prior distribution over functions to get a qualitative sense of the properties of the kernel.
In practice, this is done by choosing a grid of times $\boldsymbol{t} = \{t_i\}$, evaluating the elements of the covariance matrix
\begin{equation}
  K_{ij} = k(t_i,\,t_j;\,\hyperparams) \quad,
\end{equation}
and then generating a multivariate Gaussian sample with this covariance.
As an example, \autoref{fig:samples} shows several prior samples for three different kernel functions from \autoref{tab:kernels}, with a range of length scales $\lengthscale$ and amplitudes $\amplitude$.
In this figure, we can see some qualitative differences between the kernel functions---namely that the samples become ``smoother'' from left to right---and we can see how the characteristic input and output scales of the allowed functions change with the hyper-parameters.

A key feature of a kernel function that defines its qualitative behaviour is its mean square (MS) differentiability.
We won't formally define MS differentiability---interested readers are instead referred to Section~4.1.1 of \citet{gpml} for a more detailed discussion---but we want to highlight that MS differentiability is important because it defines the ``smoothness'' of the functions that can be modelled by the GP.
It can be seen in \autoref{fig:samples} that the exponential squared kernel function, which is infinitely MS differentiable, is significantly ``smoother'' than the exponential kernel function, which is only 1-time MS differentiable. \suz{here you say the exponential kernel is 1-time MS differentiable but in the caption of Figure 6 you say it's not MS differentiable at all. I think the caption is correct, but will let you resolve it.}

\begin{armarginnote}[]
  \entry{MS differentiability}{mean square differentiability}
\end{armarginnote}

\begin{table}[ht]
  \caption{Some kernel functions commonly used in the astrophysics literature. }
  \label{tab:kernels}
  \begin{center}
    \begin{tabular}{@{}l|l@{}}
      \hline
      Name                           &  Representation$^{\rm a}$                                    \\
      \hline
      Constant                       & $\amplitude^2$                                                     \\
      Squared Exponential$^{\rm b}$  & $e^{-(\dt/\lengthscale)^2/2}$                                              \\
      Exponential$^{\rm c}$          & $e^{-\dt/\lengthscale}$                                                  \\
      Mat\'ern-3/2                   & $\left(1 + \sqrt{3}\dt/\lengthscale\right)e^{-\sqrt{3}\dt/\lengthscale}$             \\
      Mat\'ern-5/2                   & $\left(1 + \sqrt{5}\dt/\lengthscale +5(\dt/\lengthscale)^2/3\right)e^{-\sqrt{5}\dt/\lengthscale}$  \\
      Rational quadratic             & $\left( 1 + \frac{\dt^2}{2 \gamma\lengthscale^2} \right)^{-\gamma}$ \\
      Cosine                         & $\cos 2\pi\dt/\lengthscale$                                                  \\
      Sine Squared Exponential       & $\exp\left(-\Gamma\sin^2\pi\dt/\lengthscale\right)$                      \\
      Stochastic Harmonic Oscillator$^{\rm d}$ & $\cos\left(\sqrt{1-\beta^2}\frac{\dt}{\lengthscale}\right) + \frac{\beta}{\sqrt{1-\beta^2}}\sin\left(\sqrt{1-\beta^2}\frac{\dt}{\lengthscale}\right)$                          \\
      \hline
    \end{tabular}
  \end{center}
  \begin{tabnote}
    $^{\rm a}$in each case, $\dt$ is defined as $\dt = \left|t_i - t_j\right|$, and Greek letters indicate hyper-parameters;
    $^{\rm b}$``radial basis function'';
    $^{\rm c}$``Ornstein--Uhlenbeck'', ``damped random walk'' or ``Mat\'ern-1/2'';
    $^{\rm d}$\citet{celerite}.
  \end{tabnote}
\end{table}

\begin{figure}[ht]
  \centering
  \script{samples.ipynb}
  \begin{minipage}[t]{0.3\linewidth}
    \includegraphics[width=\linewidth]{figures/samples1.pdf}
  \end{minipage} \hfill
  \begin{minipage}[t]{0.3\linewidth}
    \includegraphics[width=\linewidth]{figures/samples2.pdf}
  \end{minipage} \hfill
  \begin{minipage}[t]{0.3\linewidth}
    \includegraphics[width=\linewidth]{figures/samples3.pdf}
  \end{minipage}
  \caption{Prior samples for three different classes of kernel functions: (a) exponential, (b) Mat\'ern-3/2, and (c) exponential squared.
  In each sub-figure, the length scale $\lambda$ of the kernel increases from left to right, and the amplitude $\alpha$ increases from top to bottom.
  One thing to notice in this figure is that these kernel functions differ in their smoothness properties.
  Specifically, the exponential kernel is not mean-square differentiable, while the exponential squared kernel is infinitely differentiable.
  This can be seen qualitatively in this figure.}
  \label{fig:samples}
\end{figure}

\subsubsection{Operations on kernel functions}\label{sec:kernel-ops}

As discussed above, more flexible kernel functions are often constructed using products and sums of the standard kernels listed in \autoref{tab:kernels}.
Besides, addition and multiplication, other operations can be used to impose structure on the standard kernels.
For example, linear operations like scalar multiplication, more general affine transformations, differentiation, or integration can all be used to develop new kernel functions.

For example, there are good arguments \citep{2012MNRAS.419.3147A} to expect that the radial velocity time series of a spotted rotating star is related to the photometric time series via its time derivative.
More formally, assume we have two time series---one with observations of some function $f(t)$ and another with observations of $\dot{f}(t)$, where the dot indicates the derivative with respect to time---and the function $f(t)$ is modeled as a GP with some covariance function $k(t_i,t_j;\hyperparams)$. The covariance between an observation of $f(t)$ and one of $\dot{f}(t)$ is
\begin{equation}
  \mathrm{Cov}\left[f(t_i),\dot{f}(t_j)\right] = \left.\frac{\partial k(t_i,t;\hyperparams)}{\partial t}\right|_{t=t_j}
\end{equation}
and the covariance between two observations of $\dot{f}(t)$ is
\begin{equation}
  \label{eq:cross-deriv-covar}
  \mathrm{Cov}\left[f(t_i),\dot{f}(t_j)\right] = \left.\frac{\partial^2 k(t,t^\prime;\hyperparams)}{\partial t\,\mathrm{d}t^\prime}\right|_{t=t_i,t^\prime=t_j} \quad.
\end{equation}
\suz{I corrected a typo in the above and switched notation to use partial derivatives here since it matters which of the two inputs the derivative is taken with respect to, specially for non-stationary covariance functions.}
This means that we can compute the GP model for observations of both sets of observations and their covariances.
One technical point which is worth noting here: for \autoref{eq:cross-deriv-covar} to define a valid covariance function, the base kernel $k(t,t^\prime;\hyperparams)$ must be mean square differentiable, as mentioned briefly in the previous section.
Therefore, the exponential kernel, for example, should not be used for modeling derivative observations since it does not have the appropriate properties.

\autoref{fig:kernel-ops} illustrates the elements of GP models with different kernel functions for observations of a time series and its time derivative.
The left panels show the covariance functions for all pairwise permutations of $f(t)$ and $\dot{f}(t)$, and the right panels show samples of the function and its derivative.
The use of this type of GP kernel is discussed in more detail in \autoref{sec:RVact}, with applications to radial velocity time series observations of exoplanets.

Similarly, integration or convolution also have similar properties.
For example, real time series observations are made with finite exposure time $\Delta$, and therefore, if the underlying stochastic process $f(t)$ is modeled as a GP with covariance $k(t_i,t_j;\hyperparams)$, the observations are actually of
\begin{equation}
  f_\mathrm{int}(t) = \int_{t-\Delta/2}^{t+\Delta/2} f(t^\prime)\,\mathrm{d}t^\prime \quad,
\end{equation}
which has a covariance of
\begin{equation}
  k_\mathrm{int}(t_i,t_j;\hyperparams) = \int_{t_i-\Delta/2}^{t_i+\Delta/2}\int_{t_j^\prime-\Delta/2}^{t_j+\Delta/2} k(t,t^\prime;\hyperparams)\,\mathrm{d}t\,\mathrm{d}t^\prime \quad,
\end{equation}
which can be evaluated in closed form for some standard kernel functions.
Another context where integrated GPs have recently been used for astrophysics (albeit not for time series), is as a model of the Galactic dust distribution \citep{2022arXiv220206797M}.

\begin{figure}[ht]
  \centering
  \script{kernel-ops.ipynb}
  \includegraphics[width=0.7\linewidth]{figures/kernel-ops1.pdf}
  \includegraphics[width=0.7\linewidth]{figures/kernel-ops2.pdf}
  \caption{A demonstration of the effects of differentiation on a GP model.
  Each row shows the same figures a different base kernel: \emph{(top)} a squared exponential kernel ($\lengthscale = 1$), and \emph{(bottom)} the product of a squared exponential kernel ($\lengthscale_1 = 2.5$) with a sine squared exponential kernel ($\Gamma = 1$; $\lengthscale_2 = 3.5$).
  In each row, the leftmost panel shows the kernel function and its first and second time derivatives, which are used to evaluate the covariance matrix for a GP model for a process and its derivative.
  The right-hand panel shows a realization from the process ($f(t)$, top) and its time derivative ($\dot{f}(t)$, bottom) generated by jointly sampling from the GP including the covariance between $f(t)$ and $\dot{f}(t)$.}
  \label{fig:kernel-ops}
\end{figure}

\subsubsection{Multivariate GPs}

Since this review is focused on GPs for \emph{time domain} astronomy, we primarily consider univariate GPs with just \emph{time} as the input coordinate.
However, we would be remiss to neglect discussing multivariate GPs altogether.
Most of our discussion so far applies to multivariate datasets without any change, but there are some subtleties that are worth mentioning here.

When working with univariate GPs, it is relatively straightforward and unambiguous to compute the ``distance'' between two points $t_i$, and $t_j$: $\rho(t_i,t_j) = \dt = \left|t_i - t_j\right|$.
For multivariate inputs $\boldsymbol{x}$, more care must be taken to define a sensible distance metric $\rho(\boldsymbol{x}_i,\boldsymbol{x_j})$.
Common choices include the $d$-th norm $\rho_d(\boldsymbol{x}_i,\boldsymbol{x_j}) = \left|\left|\boldsymbol{x}_i - \boldsymbol{x}_j\right|\right|_d$ for $d=1$ or $d=2$, but other metrics---for example, the Haversine distance when the elements of $\boldsymbol{x}$ are coordinates on a sphere---might be better motivated for a particular use case.

When the input coordinates have some geometric interpretation (e.g., spatial coordinates) simple isotropic distance metrics, like those presented so far, are generally well-justified.
However, when the input coordinates are heterogeneous with, for example, different units, it is often more appropriate to use a more general distance metric.
In particular, you may want to fit for a different length scale parameter $\lengthscale_d$ for each input dimension $d$:
\begin{equation}\label{eq:ard-distance}
  \rho(\boldsymbol{x}_i,\boldsymbol{x}_j) = \sqrt{\sum_{d=1}^D \left(\frac{x_{id} - x_{jd}}{\lengthscale_d}\right)^2}
\end{equation}
or even include support for covariance between different input dimensions:
\begin{equation}
  \rho(\boldsymbol{x}_i,\boldsymbol{x}_j) = \sqrt{\boldsymbol{x}_i\,\boldsymbol{M}^{-1}\,\boldsymbol{x}_j}
\end{equation}
where $\boldsymbol{M}$ is a $D\times D$ positive definite matrix of hyper-parameters.

Kernels with a different length scale parameter for each dimension (\autoref{eq:ard-distance}) are commonly referred to as ``automatic relevance determination'' \citep[ARD;][]{gpml} kernels, the idea being that the data can inform the importance of each input dimension when describing the observations.
In practice, such a kernel is often used in conjunction with a ``shrinkage'' prior on the hyper-parameters, favouring large values of $\lengthscale_d$ as discussed further in \autoref{sec:transit_fit}.
\begin{armarginnote}[]
  \entry{ARD}{Automatic Relevance Determination}
\end{armarginnote}

A common class of multivariate GP model for time domain astronomy are datasets with multiple parallel covariant time series.
For example, multi-band time series produced by surveys like PanSTARRS or LSST, or radial velocity time series with parallel activity indicators.
One useful way to specify these datasets are to define the inputs as $\boldsymbol{x_i} \equiv (t_i,\ell_i)$ where $t_i$ the time of the $i$-th observation and $\ell_i$ is a ``label'' for which time series the $i$-th observation is drawn from.
In this case, one common choice of kernel function is:
\begin{equation}\label{eq:multiband}
  k(\boldsymbol{x}_i,\boldsymbol{x}_j;\hyperparams) = \boldsymbol{a}_{\ell_i}^\mathrm{T} \boldsymbol{a}_{\ell_j} k_0(t_i,t_j;\hyperparams) \quad,
\end{equation}
where $k_0(t_i,t_j;\hyperparams)$ is a standard one-dimensional kernel, and the set of $\{\boldsymbol{a}_d\}_{d=1}^D$ are also hyper-parameters of the model.
This class of kernels includes the models commonly used for interpolating multi-band photometric observations of transients \citep[e.g.,][]{2020ApJ...905...94V, 2021ApJS..255...24V}.
In the case where the data are rectangular (i.e., all input dimensions are observed at all times), the covariance matrix defined by \autoref{eq:multiband} will have Kronecker structure \citep[e.g.,][]{2015arXiv151101870W, 2020AJ....160..240G}, which can be exploited to improve the computational performance of the GP model.

% \dan{More here.} \suz{There's examples of multi-variate GPs where time is only one of several inputs applied to modelling systematics in transit observations in \ref{sec:transit_fit} you can refer to here if you like. In particular I mention the ARD kernel there.}

\subsubsection{Physically motivated kernels}

Physical processes that are expected to be GPs appear frequently in astrophysics.
For example, a physical model that is specified by a parametric form for its ``power spectrum'' in the Fourier domain (think, the cosmic microwave background, or solar-like asteroseismic oscillations), could \emph{equivalently} be specified by its covariance function (or kernel function) in real space.
It is common practice in astrophysics to analyse datasets with these models by first Fourier transforming the data, and then fitting these transformed data using a physical model for the power spectrum.
There is a rich literature on this topic, discussing the motivation for this approach, and the rigorous propagation of measurement uncertainty for these datasets.
These analyses could also be performed using GPR without first transforming the data, but this has typically been avoided since na\"ive implementation of GPR would be significantly more computationally expensive, or even completely intractable.

There are, however, many cases where analysis in the time domain is preferred---at least in principle---over a Fourier analysis.
This is especially true when working with short, significantly unevenly sampled, or otherwise heterogeneous datasets, or when combined with non-stationary functions that cannot be compactly represented in the frequency domain. 
% \dan{More here or cut?} \suz{I think it's worth making the general point that in principle it's always better to work in the time domain when the sampling is irregular, rather than interpolate in order to compute a power spectrum? I'm also uncertain how well defined FTs of heteroskedastic datasets are?}

Another class of GPs that are commonly used for astrophysics are stochastic differential equations \citep[see][for a detailed discussion of these models]{Sarkka:2019}, although this terminology is not standard in the literature.
The most widely-used model in this class is the Ornstein--Uhlenbeck \citep{1930PhRv...36..823U}, or ``damped random walk'' process \citep[e.g.,][]{2009ApJ...698..895K, 2012A&A...546A..89B, 2020AJ....160..265H}.
These models---or the generalized ``continuous-time autoregressive moving average'' (CARMA) process \citep[e.g.,][]{2014ApJ...788...33K, celerite, 2022arXiv220108943Y}---are commonly used as a physically-motivated model for the stochastic time-variability of active galactic nuclei (AGN).
Although this fact is not always noted in the literature, these models are examples of GPs with specific kernel function choices.
Besides their physical interpretability, these models can also be evaluated with computationally efficient and scalable algorithms \citep[e.g.,][]{2014ApJ...788...33K, celerite}, which can be crucial for practical applications, as we discuss in \autoref{sec:fast}.
% \dan{Do we want to add some of the math here? It might be hard to make it useful without too many details.} \suz{If there's a didactic reference we can cite it might be better to do that. NB might want to cite the work of Caceres, Feigelson et al. using ARMA models for transit searches? \url{https://ui.adsabs.harvard.edu/abs/2019AJ....158...58C/abstract}}

Asteroseismology is another domain where Fourier space models are common, and recently there has been some work using time-domain GPR to analyse these datasets \citep[e.g.,][]{2017AJ....154..254G, 2018ApJ...865L..20F}.
The kernel functions used for these analyses are typically parameterized by the frequency and amplitude of the oscillation modes, but the likelihood is evaluated using \autoref{eq:gp_like} in the time domain.
Time-domain GP models for asteroseismic oscillations can be particularly useful when combined with a mean model that is compactly specified in the time domain \citep[e.g., a transiting planet,][]{2017AJ....154..254G}, or when the time sampling of the data is strongly non-uniform with complicated windowing effects \citep{2018ApJ...865L..20F}.
Like the AGN models discussed above, these asteroseismic GP models can typically be evaluated efficiently using scalable algorithms.
So far, the applications of GPR to asteroseismology have been generally restricted to simple effective models with a single wide mode used to capture the power spectrum of the oscillations, but it is conceivable that more complex models could be applied to fully capture the structure of the power spectrum.

While the models discussed so far in this section have strong physical motivations, it is also common to apply GPR using effective kernel models with physically-motived \emph{parameters}.
For example, GPs have been used to measure the rotation periods of stars in photometric time series using a quasi-periodic kernel function, where one parameter of the model can be interpreted as the stellar rotation period \citep{2018MNRAS.474.2094A, 2020MNRAS.492.1008G}.
These models perform well and can be used to formally propagate the observational uncertainty to the inferred rotation period.
Recently, some more physically interpretable kernel functions have been developed for measuring stellar rotation, and mapping stellar surfaces using photometric time series \citep{2021AJ....162..124L}, but these models have so far seen limited use because of their computational cost.

% \suz{Mention kernels with break points to detect sudden changes, as described in review by \citet{2012RSPTA.37110550R} for example?}

\subsection{Hyperparameter Inference}\label{sec:inference}

A key component of all the covariance functions discussed above is that they are all \emph{parameterized} by a set of hyper-parameters.
In most cases, you won't have \emph{a priori} knowledge for how the values of these hyper-parameters should be set.
Instead, their values will need to be numerically tuned or incorporated into a larger inference scheme.

In the astrophysics literature, the most common approach for taking this uncertainty into account---and the method that we advocate for here---is to treat the hyper-parameters directly as parameters of the model.
In other words, instead of just fitting for the parameters of the mean model $\meanparams$, we can simultaneously fit for both $\meanparams$ and the hyper-parameters $\hyperparams$.
In \autoref{sec:gp-lsq-link}, we defined the likelihood for a GP model and, in \autoref{sec:gp-inf}, we sketched the procedure used for Bayesian inference with such a model.
The likelihood function defined in \autoref{eq:gp_like} is a function of both $\meanparams$ and $\hyperparams$, and we can use that function as an objective for a non-linear optimization routine to find the maximum likelihood parameter values, or in a Markov chain Monte Carlo (MCMC) procedure to marginalise over the hyper-parameters, and propagate their uncertainty to constraints on the parameters of the mean model.

An important point here is that, since many data analysis procedures in astrophysics include a step like the ones listed above, the use of a GP likelihood doesn't significantly change the processing.
In fact, we like to say that the GP likelihood can be used as a drop-in replacement for anywhere you're currently using a ``chi-squared'' objective.
There are some practical reasons why things aren't necessarily this simple (for example, computational cost, as described below), but the sentiment stands.

In this review, we won't go into too many details about the inference algorithms, but throughout the text, we will regularly use the \textsf{BFGS} gradient-based, non-linear optimization routine \citep{Nocedal:1999, scipy} to find the maximum likelihood parameter values
\begin{equation}
  \meanparams_\star,\hyperparams_\star = \argmax_{\meanparams,\hyperparams} \,\mathcal{L}(\meanparams,\hyperparams) = \argmax_{\meanparams,\hyperparams} \,\log p(\ydata\,|\,\meanparams,\hyperparams) \quad.
\end{equation}
Another common inference technique used in astrophysics---and in this review!---is the use of Markov chain Monte Carlo (MCMC) to generate posterior samples
\begin{equation}
  \meanparams,\hyperparams \sim p(\meanparams,\hyperparams\,|\,\ydata)
\end{equation}
that can be used to marginalise over some subset of the parameters, and estimate the uncertainty on the parameter values.
All the examples in this review are implemented using the \textsf{NumPyro} probabilistic inference library, and all the MCMC examples use the gradient-based No U-Turn Sampler (NUTS) algorithm \citep{Hoffman:2014}.

\autoref{sec:quasar} and \autoref{sec:transit}, and specifically \autoref{fig:quasar} and \autoref{fig:transit}, show the results of worked examples of an MCMC-based GPR workflow, without and with (respectively) a non-trival mean function.

\subsection{Model assessment, validation, and selection}
\label{sec:model_selection}

Given the wide array of possible kernel functions described in \autoref{sec:kernels}, it can be important to assess the performance of your model and the relevant choices.
This includes both assessing the choice to use a GP in the first place, and the specific choice of kernel function.
It's important to note that there's nothing fundamentally different about GPs in this context when compared to other models for data, so that means that many methods that you may already use for model selection and validation in other contexts will also apply when using GPs.
That being said, within the astrophysics literature formal probabilistic model checking has had limited use, and GPs do come with some specific technical complications, therefore we will discuss some examples of model validation and selection techniques that have been used for GPs.

\dan{Say something here about assessing Gaussianity, and more generally model checking (rather than comparison; posterior predictive checks).}

When it comes to selecting between different possible kernel functions, the approach that you take may depend on your specific research goals.
For example, in many cases, including the transiting exoplanet example in \autoref{sec:transit}, the main parameters of physical interest may be the parameters of the mean model, and the GP is simply an effective model for stochastic nuisances.
In this case, it may be sufficient to demonstrate that the inferred results are not significantly inconsistent for different choices of kernel function.

Other common use cases, like the time delay example in \autoref{sec:quasar}, primarily require good predictive performance for the GP model.
In these cases, methods like cross-validation \citep[e.g.,][]{gelfand1992model} or posterior predictive assessment \citep[e.g.,][]{gelman1996posterior} can be used to evaluate different choices of kernel function.

To demonstrate these approaches, \autoref{fig:assessment} shows the results of performing a model comparison between three different kernel functions applied to a simulated dataset.
The simulated data were generated from a GP model with a squared exponential kernel with known parameters, and we aim to compare the performance of three kernels: (1) the squared exponential kernel, (2) the Mat\'ern-3/2 kernel, and (3) the rational quadratic kernel.
The simulated dataset is shown in the left panel of \autoref{fig:assessment}.

First, using the full dataset, we compute the Bayesian evidence integral for each of these model choices using a nested sampling algorithm implemented in the \textsf{jaxns} package \citep{jaxns}.
The Bayesian evidence integral is defined as:
\begin{equation}
  Z(H) \equiv p(\ydata\,|\,H) = \int p(\meanparams,\hyperparams\,|\,H)\,p(\ydata\,|\,\meanparams,\hyperparams,H)\,\mathrm{d}\meanparams\,\mathrm{d}\hyperparams
\end{equation}
where $H$ indicates the modeling choices, in this case the choice of kernel function.
For better or worse\footnote{A discussion of the relative merits of different model selection algorithms is beyond the scope of this review. \suz{Sure but can we link to a more in-depth discussion of the topic for interested readers?}}, the Bayesian evidence is frequently used in the astrophysics literature as an ingredient in model selection procedures.
Nested sampling is an algorithm for numerically estimating this integral, and we apply it to produce estimates of $Z$ for each choice of kernel $H$, and plot these results in the middle panel of \autoref{fig:assessment}.
In this figure, it is clear that the squared exponential and rational quadratic kernels are indistinguishable under this metric, while the Mat\'ern-3/2 kernel is somewhat disfavoured.

Another popular method for model selection that is less commonly used in the astrophysics literature is cross validation.
Cross validation is designed to assess the predictive performance of the model, and it proceeds by holding out some data, fitting the rest of the data, and then computing the likelihood of the held out data conditioned on the fit results.
These steps can then be repeated for different held out samples.
For a GP model, the likelihood of the held out data can be computed using the predictive distribution discussed in \autoref{sec:pred}.
In particular, the likelihood is the following multivariate Gaussian
\begin{equation}\label{eq:held-out-like}
  p(\ydata_\star\,|\,\ydata,\meanparams, \hyperparams) = \mathcal{N}(\boldsymbol{f}_\star,\boldsymbol{C}_\star)
\end{equation}
where $\boldsymbol{f}_\star$ and $\boldsymbol{C}_\star$ are defined in \autoref{eq:pred}, noting (importantly!) that the (squared) observational uncertainties on the held out data should be included on the diagonal of $\boldsymbol{C}_\star$.

In this example, we use MCMC to fit the data shown as black dots in the left panel of \autoref{fig:assessment}, holding out the data points indicated by blue crosses.
At each step in the MCMC, we evaluate the likelihood in \autoref{eq:held-out-like} for the held out data conditioned on the training data and the model parameters.
In the rightmost panel of \autoref{fig:assessment}, we plot the posterior distribution of the held out log probability to show that, like with the evidence integral, the squared exponential and rational quadratic kernels are indistinguishable, while the Mat\'ern-3/2 kernel is disfavoured.
In this case we hold out a single fixed subset of the data, but it is common practice to repeat this process for several unique held out subsets, and average these results.
There is some freedom in the choice of held out data, and it's not trivial to select the best procedure for correlated time series data.
In this example, we held out some contiguous data points at the beginning and end of the time series to assess the model's performance when extrapolating, and then held out a random subset of the remaining data points to validate the bulk statistics of the process.
For GP models, simple ``leave-one-out'' or other uniform sampling schemes are unlikely to be sufficient for assessing the correlation structure of the data, and care should be taken to select validation datasets that capture the details of interest.

\begin{figure}[ht]
  \centering
  \script{assessment.ipynb}
  \includegraphics[width=\linewidth]{figures/assessment.pdf}
  \caption{A demonstration of model assessment using data simulated from a GP model using a squared exponential kernel with known parameters.
  \emph{(left)} The data simulated data are shown as black dots, with some held out for cross validation indicated by blue crosses.
  The maximum likelihood estimate of the GP's predictive distribution based only on the black points is indicated with the orange $1-\sigma$ contours.
  \emph{(middle)} The value of the Bayesian evidence integral numerically estimated using nested sampling and all the data in the left panel (including the data held out for cross-validation), is shown for each kernel function.
  \emph{(right)} The posterior distribution of the held out log probability is plotted for each kernel function.}
  \label{fig:assessment}
\end{figure}

% Heading 3
\section{GAUSSIAN PROCESSES IN TIME-DOMAIN ASTRONOMY}
\label{sec:uses}

In this section we discuss a range of applications of GPR to time-domain datasets in the astronomical literature. The present-day popularity of GPs in astronomy precludes any attempt at an exhaustive review. Instead, we have selected examples which showcase the power and flexibility of the method. No special meaning should be ascribed to the order in which these applications are discussed, beyond the fact that we started with the application domains we are most familiar with.

\subsection{Stars and exoplanets}
\label{sec:stars_planets}

It is no coincidence that the recent uptake of GPR in astronomy coincides with the meteoritic rise of exoplanet studies. Almost all exoplanet detections and related discoveries are made indirectly, and involve detecting small and/or short-lived signals in time-domain datasets: transits in light curves, Keplerian signals in Radial Velocity (RV) or astrometry time-series. These are invariably affected by, and often buried in, correlated noise or nuisance signals, making GPR a natural choice.

%The search for exoplanets has also motivated the development of large-scale time-domain photometric and spectroscopic surveys, a by-product of which is the study of a wide range of stellar astrophysics phenomena, where GPR has also found fresh applications that will be discussed later in this Section. Historically, the first applications of GPR to exoplanet datasets were for transit modelling.

\subsubsection{Transit searches}
\label{sec:transit_det}

Following the discovery of the first transiting exoplanet \citep{2000ApJ...529L..41H,2000ApJ...529L..45C}, numerous ground-based photometric monitoring surveys were set up (or re-purposed) to search for planetary transits. Early estimates of their yield \citep[see e.g.][]{2003ASPC..294..361H} ran into the hundreds of planets per year, but these soon turned out to be highly optimistic. While surveys such as the Optical Gravititational Lensing Experiment (OGLE) did indeed discover numerous transit-like events
%\citep{2002AcA....52....1U,2002AcA....52..317U,2003AcA....53..133U,2004AcA....54..313U}
\citep[see e.g.][]{2002AcA....52....1U}, only a handful were ultimately confirmed as bona-fide transiting planets, and the process took years. %\citep{2003Natur.421..507K,2003ApJ...597.1076K,2004ApJ...609L..37K,2005ApJ...624..372K,2004A&A...421L..13B,2004A&A...426L..15P}.
The vast majority of early transit candidates were later diagnosed as diluted or grazing stellar eclipses based on radial velocity follow-up.
%\citep{2005A&A...431.1105B,2005A&A...438.1123P}.
Intriguingly, a small fraction turned out to be spurious detections, despite having relatively high signal-to-noise ratio (SNR). Indeed, \citet{2006MNRAS.373..231P} noted that all the confirmed OGLE planets had very high SNRs, and went on to show that the SNRs for all candidates had been over-estimated due to the presence of correlated, or red, noise in the light curves. They proposed an empirical method to evaluate the red noise on transit time-scales from the individual light curves (analogous to the method proposed by \citealt{prh92a} to evaluate the correlation matrix of a quasar light curve), and then a prescription for adjusting the detection threshold accordingly. In effect, they were searching for a way to model the covariance of the data. This is precisely what GPR allows, but the methodology was not known in the exoplanet community at the time.

GP models have been used on occasion to \textit{detrend} light curves from transit surveys prior to running transit searches, i.e.\ to remove variability on timescales significantly longer than a transit \citep[see e.g.][]{2016ApJS..226....7C}, and are used frequently after detection to model the out-of-transit baseline alongside the transit signal itself (see Section~\ref{sec:transit_fit}). In principle, a GP could also be used to model out-of-transit variations (intrinsic or instrumental) alongside transits as part of a detection pipeline. Doing this simultaneously rather than sequentially allows more flexible models to be used for the out-of-transit variations, and avoids the risk of corrupting the transit signal by using an excessively aggressive filter at the detection stage, as demonstrated by \citet{2015ApJ...806..215F} for \textit{K2} data. In that case, the primary source of out-of-transit variability is instrumental systematics (see Section~\ref{sec:transit_fit}), which are to some degree common to all light curves in a given observing run. \citet{2015ApJ...806..215F} model these systematics as a linear combination of the first 150 principal components of the ensemble of light curves, alongside a simple box-shaped transit model, varying the latter's period, phase and duration.

In other space-based transit surveys such as \textit{CoRoT}, \emph{Kepler} and \emph{TESS}, the systematics are less prominent, so the dominant source of out-of-transit variations is intrinsic stellar variability, which occurs mainly on timescales longer than transits, and can therefore be filtered out quite effectively without removing the transits. However, in some cases, for example for rapidly rotating and magnetically active stars, the separation between transits and stellar signal becomes less clean, significantly reducing the performance of sequential detrending approaches (irrespective of the algorithm used, \citealt{wotan}).
In such cases, a simultaneous modelling approach could improve detection performance significantly. Since the nuisance signal is specific to each star, it cannot be modeled using other light curves, but a GP is a credible alternative. Such an approach could also offer a small but significant enhancement in the sensitivity future missions such as \textit{PLATO} to transits of habitable planets around Sun-like stars (which are very shallow and relatively long). The main issue would be computational cost: transit surveys monitor $10^4$--$10^5$ stars at a time, with $10^3$--$10^5$ observations per run, and the transit search must be run over a fine grid of periods and phases. Doing this with standard GPR methods would be impractical. However, now fast and scalable GP solvers are available, this becomes a more feasible proposition.

\subsubsection{Transit and eclipse modelling}
\label{sec:transit_fit}

In practice, GPR has been much more widely applied to transit modelling than to detection \textit{per se}. The detailed depth, timing, duration and shape of a transit depends on the planet-to-star radius ratio $R_{\rm p}/R_\star$, the system scale $a/R_\star$ (where $a$ is the orbital semi-major axis), the impact parameter $b$, the orbital period $P$ and the time of transit centre, $T_0$ as well as the limb-darkening profile of the star. High-precision observations of one or more transits of a given planet can be used to infer these parameters, which has a wide range of scientific applications. The most obvious is to estimate the planet size (and hence, given a mass estimate, its bulk density and composition). The system scale is directly related to the stellar density \citep{2003ApJ...585.1038S}, which can be used for sanity checks to ensure the transits are indeed of planetary origin [ref tingley]. Small departures from strictly periodic timing, known as Transit Timing Variations (TTVs), probe dynamical interactions between multiple planets in a given system, and can be used measure the planets' masses and eccentricities and to reveal the presence of additional, non-transiting planets \citep{2005Sci...307.1288H}. Wavelength-dependent measurements of the transit depth probe the effect of the planet's atmosphere on the star light that filters through it, and hence allow us to access the atmospheric composition \citep{2000ApJ...537..916S}. The depth of any secondary eclipse (when the planet passes behind the star) depends on the planet-to-star flux ratio, and can therefore be used to measure an emission or reflection spectrum, while its timing yields strong constraints on the orbital eccentricity. All of this requires precise and accurate measurements of the transit parameters, for which any correlated noise must be accounted for explicitly, as we demonstrated using a simulated example in Section~\ref{sec:transit}.

To address this, \citet{2009ApJ...704...51C} proposed a wavelet-based method to model correlated noise in transit light curves, which can be seen as a special case of GPR with a covariance function belonging to the exponential family, but formulated in terms of wavelets. Rather than modelling the covariance in the time domain, the Power Spectral Density (PSD) of the noise is assumed to be of the form $P(f) \propto f^{-\gamma}$.
Noting that the wavelet transform of such a process gives rise to a nearly diagonal covariance matrix, \citet{2009ApJ...704...51C}  derived an expression for the likelihood that can be computed in $\mathcal{O}(N)$ operations. 
This method has been used widely since its publication to model transit observations from space missions which stare continuously at a given field, such as \textit{Kepler} or \textit{TESS}. \begin{armarginnote}
  \entry{Colours of noise}{White, pink and red noise correspond to $\gamma=0$, $1$ and $2$, respectively.}
\end{armarginnote}
However, the use of a wavelet transform requires regular time sampling, which precludes its application to datasets with irregular sampling or significant data gaps, for example from space telescopes in low-Earth orbit such as Hubble and Spitzer.

This limitation is significant, as Hubble and Spitzer have been the workhorse instruments for transit and eclipse spectroscopy for over a decade. This involves measuring minute changes in the depth of the primary transit or secondary eclipse as a function of wavelength, either using successive photometric observations through different filters or using a spectrograph (see \citealt {2018haex.bookE.100K} for a review). The signal of interest is the wavelength dependence of the transit depth, or the depth of the eclipse, which are of order $10^{-4}$ and $10^{-3}$ respectively, in  the most favourable cases. Even when observed from space, these signals are typically dwarfed by instrumental systematics. The pointing jitter and thermal relaxation of the telescope as it orbits the Earth causes the target star to move on the detector, typically by a fraction of a pixel. As the sensitivity of the detector varies from pixel to pixel (the ``flat-field'') and between the centre and edges of each pixel, this motion causes spurious variations in the recorded flux from the target. Although we expect the measured flux to depend on ``housekeeping" variables such as the satellite orbital phase, telescope attitude, the centroid of the image or the locus and angle of the spectrum of the detector, or the temperatures of various parts of the instrument, a physically motivated model for the form of this dependence is generally lacking. \textit{Ad hoc} parametric (e.g.\ polynomial) models are problematic, as the choice of model inputs and functional form is arbitrary, yet can drastically alter the resulting exoplanet spectrum \citep{2011MNRAS.411.2199G}.

To address this, \citet{2012MNRAS.419.2683G} proposed a GPR framework to model the systematics in space-based transit observations. The aforementioned housekeeping variables are treated as multi-dimensional inputs to a squared exponential GP whose mean is the transit signal, allowing one to marginalise over broad families of systematic models without assuming a specific functional form for them, all the while propagating the resulting uncertainties on the physical parameter of interest, namely the (wavelength-dependent) planet-to-star radius ratio. The kernel used was a squared exponential kernel with an automatic relevance determination (ARD) distance metric (\autoref{eq:ard-distance}), this allows one to try including a wide range of housekeeping data into the fit; in principle only those which are genuinely relevant will have an effect on the result.
GP-based systematics models of this kind were later extended to eclipse spectroscopy, where the parameter of interest is the planet-to-star flux ratio, which controls the eclipse depth. Among other successes, this approach has enabled the first measurement of the wavelength-dependent albedo of a hot Jupiter \citep{2013ApJ...772L..16E}, helped resolve early controversies surrounding the treatment of systematics in Spitzer observations \citep{2015MNRAS.451..680E} and
led to the first unambiguous detection of a thermal inversion in an exoplanet emission spectrum \citep{2017Natur.548...58E}. Using simulated data, \citet{2014MNRAS.445.3401G} showed that GPs outperform parametric models when the true form of the systematics is unknown, but also that the most robust results overall are obtained by marginalising over families of both parametric and GP models.

Despite these theoretical advantages and practical successes, the use of GPR to analyse low-resolution transit and eclipse spectra remains confined to a relatively small subset of the corresponding community. Parametric models (in some cases marginalising over families thereof, \citealt{2016ApJ...819...10W}) remain the most widely-used approach for Hubble space telescope observations, and Pixel-Level-Decorrelation (PLD) for Spitzer observations \citep{2015ApJ...805..132D}. Possible explanations for this include computational cost, as well as the fact that these other methods are perceived as easier to implement and interpret. As the recently-launched JWST begins to deliver much higher signal-to-noise observations of exoplanet transits \citep[see e.g.][]{https://doi.org/10.48550/arxiv.2208.11692}, with considerably lower systematics, it will be interesting to see how this topic evolves in the next few years.

One general shortcoming of current GP-based systematics models with multi-dimensional inputs is that the uncertainties on the input variables are ignored. This could in principle be remedied by treating the housekeeping variables as noisy observations of latent GP variables, and modelling them alongside the observed fluxes, but we are not aware of published attempts to do this in practice to date. \begin{armarginnote}
  \entry{Latent variable}{A random variable which is never actually observed.}
\end{armarginnote}
In spectroscopic observations, whether using GPs or parametric models, it has become standard practice to model the ``white" light curve first (obtained by integrating the spectrum over the full wavelength range at each time-step). ``Coloured'' light curves, extracted in individual wavelength bins, are then divided by the best-fit systematics model derived from the white light curve, before being modelled further. To our knowledge, no attempt has been made to model the wavelength dependence of the systematics directly.

On the other hand, GPR has become quite widely used for modelling correlated noise in single-band transit observations. One striking example is the case of transiting planet candidates discovered by \textit{Kepler} around giant stars. By comparing the stellar densities derived from transit modelling to those expected from independent estimates, \citet{2014ApJ...788..148S} argued that many of these candidates, including the confirmed planet Kepler-91, might be false positives. However, red giant light curves contain stochastic variability on timescales of hours due to granulation. In the case of Kepler-91, \citet{2015ApJ...800...46B} showed that the apparent density discrepancy disappears when this granulation signal is modelled using a GP.

\subsubsection{Systematics removal in the presence of stellar variability}
GP-based models have also been used to correct instrumental systematics in data from the \textit{K2} space mission, which used the \textit{Kepler} satellite to perform an Ecliptic plane survey after the failure of two of its reaction wheels. During the \textit{K2} observations, the satellite underwent significant roll-angle variations, causing the stars to move on the detector by more than a pixel over a timescale of hours. The satellite thrusters were fired every $\sim 6$\,h to return the spacecraft to its nominal attitude, but the resulting drift caused significant changes in the measured stellar fluxes, due to the detector inter- and intra-pixel variations, and to changes in the contamination of the photometric aperture by neighbouring targets as well as in aperture losses. These can be modelled effectively using a GP with a squared exponential covariance function depending on the roll angle \citep{2015MNRAS.447.2880A, 2016ApJS..226....7C} or the star's 2-D position on the detector \citet{2016MNRAS.459.2408A}. In these approaches, a second, time-dependent term is added to the GP covariance function to represent the target star's intrinsic variability. This not only improves the fit but allows the position-dependent systematics to be evaluated separately and thus removed while preserving intrinsic variability. The use of a time-depedent GP term improves the photometric performance over other widely-used methods for correcting \textit{K2} systematics \citep[e.g.,][]{2014PASP..126..948V} when the stellar variability is significant and/or occurs on timescale similar to the roll angle variations. Ultimately, the best overall photometric precision for \textit{K2} was achieved by combining PLD to model the position-dependent systematics with a time-dependent GP to model intrinsic variability \citep{everest1,everest2}.

\begin{textbox}[ht]
\section{Composite GPs}
A composite (additive or multiplicative) GP is a powerful way to separate different components of a time-series dataset, whether these components depend on different input variables (as in the instrumental systematics examples discussed in this section) or not. To evaluate the predictive distribution for a particular component of a composite GP, one simply evaluates $\mathbf{K}_\star$ and $\mathbf{K}_{\star\star}$ in Equations~(\ref{eq:pred}) using that component only, while using the full covariance for $\mathbf{K}$.
\end{textbox}

\subsubsection{Quasi-periodic GP models for stellar light curves}

The light curves of Sun-like stars (broadly construed) display low-amplitude quasi-periodic variations, which are caused by the rotational modulation and evolution of magnetically active regions on their surfaces. These produce quasi-periodic variations in both photometry and radial velocity observations, and GPs have in recent years become one of the most popular ways of modelling them. One reason for this is that a simple quasi-periodic covariance function (Equation~\ref{eq:kqp} gives rise to functions, which reproduce the light curves of rotating stars with evolving active regions remarkably well.

%\suz{Some of this paragraph may be superseded by text in Section 3.} In particular, the following quasi-periodic (QP) kernel has proved especially popular to model stellar activity signals:
%\begin{equation}
%    k_{\rm QP} = A \exp\left[ - \Gamma \sin(\pi \tau / P) - \tau^2 / \lambda \right],
%\end{equation}
%where $A$ is the overall variance, $P$ is the period and $\Gamma$ controls the harmonic complexity of the periodic term, while $\lambda$ controls the evolution timescale. This is only one of many possible options: a valid quasi-periodic kernel can be constructed by multiplying together any valid periodic covariance function with any other valid covariance function that gives rise to a decaying envelope, and a valid periodic covariance function by replacing the argument(s) of an aperiodic one by a periodic function thereof. In this particular example, the periodic term $\exp[-\Gamma\sin^2(\pi t/P)]$ is obtained by considering a squared exponential in 2-D space $(u,v)$ and setting $u=\cos\tau$ and $v=\sin\tau$ [REF Mackay].

GP models were first applied to the simulated observations of the Sun's total irradiance variations by \citet{2012MNRAS.419.3147A}, to test a new method to simulate Radial Velocity (RV) variations based space-based photometry on solar data. The context of this work was the large numbers of candidate transiting exoplanets being discovered at the time by the \textit{CoRoT} and \textit{Kepler} space missions. Detecting the planets signals in RV was needed to confirm the planetary nature of the candidates by measuring their masses, but was being hampered by the apparent RV variations caused by active regions. Using simple geometric considerations, \citet{2012MNRAS.419.3147A} derived a simple (approximate) relationship between the flux perturbation caused by active regions, $F$, and their RV signature, which depends on both $F$ and its time-derivative $F'$. They tested this data-driven ``$FF'$ method'' on simulated photometric and RV observations of the Sun-as-a-star produced using resolved magnetograms, using a GP as a principled smoothing tool to evaluate $F$ and $F'$ from the photometry, and comparing the results to the simulated RVs. \citet{2012MNRAS.419.3147A}  tested a number of covariance functions, both aperiodic and Quasi-Periodic (QP), and found the former to be preferred in the solar case, as expected if most active regions on the Sun do not persis for much longer than its rotation period. More active and/or more rapidly rotating stars display variability that is coherent over multiple rotation periods, making QP GPs the model of choice. For example, \citet{2014MNRAS.443.2517H} used a QP kernel when applying the aforementioned $FF'$ method to the active planet-host star CoRoT-7.

Today, using GPs to model light curves containing stellar variability is standard practice. The kernel most frequently used for this purpose is the aforementioned QP kernel, whose simplicity and flexibility make it a popular choice. \citet{2018MNRAS.474.2094A} implemented a Bayesian inference framework based on this kernel to measure accurate rotation periods from \textit{Kepler} light curves. The very complex dependence of the likelihood surface on the parameters, particularly the period, makes this challenging, and careful tuning of the posterior sampling strategy is required.

Although the QP kernel provides a phenomenological rather that physically motivated description of the variability, some of its parameters lend themselves to a physical interpretation in terms of the star's rotation period and the evolution timescale of active regions. \citet{2022MNRAS.tmp.2007N} tested this using simulated light and RV curves based on physical star-spot models and confirm that, for moderately well-sampled datasets, the period of the GP does indeed provide a precise and accurate measure of the stellar rotation period. The same is true, albeit to a lesser extent, for the evolution timescale: the correlation between simulated and recovered values is more scattered, and breaks down when the datasets spans less than the simulated evolution timescale.

On the other hand, the QP kernel (or any covariance function of time only) does not give access to the physical properties of the spots, such as their size, latitude or contrast. The problem of inferring these properties from light curves is fundamentally ill-posed \citep{2021AJ....162..123L}, making direct inference of individual spot properties or brightness maps highly degenerate (whatever the methodology used). However, \citet{2021AJ....162..124L} derive a closed-form expression for a GP that describes the light curve of a rotating, evolving stellar surface conditioned on a given distribution of starspot sizes, contrasts, and latitudes. This can be used in a hierarchical Bayesian framework to infer the distribution in question from ensembles of light curves.

Most studies focusing on the photometric signatures of stellar activity ignore variations on short timescales, either working with binned data or using a ``jitter term'' to absorb them. This keeps the model simple and, when binning, speeds up computing time. However, when the time-sampling and precision of the observations allow it (e.g.\ for data from \textit{CHEOPS} or \textit{PLATO}), composite GP models with different terms to describe activity, granulation and stellar oscillations can be useful. For example, \citet{2020A&A...634A..75B} analysed light curves of stars observed at high cadence and high precision in the \textit{CoRoT} asteroseismology field. They showed that the parameters of planetary transits injected into these light curves were recovered more accurately when using a composite GP to model the stellar variability than one with a single term. They also tested a white noise only model, which provided less precise precise estimates than either GP models, albeit generally consistent with both.

\subsubsection{Stellar activity in Radial Velocities} \label{sec:RVact}
Given the remarkable instrumental precision actived by modern RV spectrographs, stellar activity is nowadays the key factor limiting the sensitivity of RV surveys low-amplitude and/or long-period planet signals. The effect of active regions on RV observations is two-fold. First, as dark spots rotate on the stellar surface, they distort the profile of spectral lines, removing a small contribution first from the red wing of each spectral line, then from the line core, and then from the blue wing. The second effect is more subtle: in the absence of active regions, spectral lines of Sun-like stars display a net blue-shift due to granulation: the emission from the hot, up-welling material in the granules, which is blue-shifted, dominates over that from the cooler material falling back down in the inter-granular lanes). In a facula, i.e.\ a region of enhanced magnetic flux density compared to the ``clean'' photosphere, convection is suppressed, leading to a localised reduction in this ``convective blue-shift''. Faculae have very small photometric contrast but can cover a much larger area than dark spots, so their RV signature can dominate over that of spots, specially for moderately slow rotators like the Sun with modest magnetic fields \citep{2010A&A...512A..39M}.

\begin{figure}[ht]
  \includegraphics[width=\linewidth]{figures/RV_framework.pdf}
  \caption{Schematic of the multi-dimensional GP framework of \citet{2015MNRAS.452.2269R} for modelling activity signals in RV data. The observed RVs (shown as orange points with error bars in the bottom right panel) are modelled alongside observations of activity indicators extracted from the spectra or line profiles (orange points with error bars in the bottom left and bottom middle panels). In all these time-series, the activity signal (shown by the black line in each panel) is modelled as a linear combination of an unobserved (latent) GP, shown in the top-left panel, and its first time-derivative, shown in the top middle panel. In addition, the Keplerian signal(s) of putative planet(s) (the sum of which is shown by the blue lines in the top right and bottom right panels) contribute to the RVs but not to the activity indicators. This hierarchical model is implemented in a Bayesian framework, sampling over all the hyper-parameters listed in the blue boxes (subject to appropriate priors) to derive posteriors over the hyper-parameters of interest and/or estimate the evidence for models with different numbers of planets. }
  \label{fig:RV_framework}
\end{figure}

Overall, active regions produce quasi-periodic RV variations which are frequently modelled with the same kind of QP GPs as for light curves. However, unlike transits, planetary signatures in RV occur on timescales similar to activity signals, meaning that the risk of overfitting is significant. This is partly mitigated for RV follow-up of transiting planets by strong priors on the stellar rotation period and on the period and phase of the planetary orbit, but it is a severe problem for ``blind'' RV searches. The issue is compounded by the fact that RV observations are invariably ground-based, and often have sparse timee-sampling.

As previously mentioned, \citet{2012MNRAS.419.3147A} showed that an approximate relationship should exist between the photometric and RV signatures of active regions. Specifically, if $F(t)$ is the photometric signature and $F'(t)$ its time-derivative, the RV variations are expected to scale as $A F(t) F'(t) + B F^2(t)$, where the first term corresponds to spots and the second to faculae (assuming that the two are generally co-located, although the latter covers a larger area), and $A$ and $B$ are tunable free parameters. In its original form, however, this method was of limited applicability, as it requires the RV observations to be quasi-simultaneous with high-precision, tightly sampled light curves, which the case only exceptionally (see e.g.\ \citealt{2014MNRAS.443.2517H}).

On the other hand, RV extraction pipelines routinely provide additional indicators measuring the characteristic width and asymmetry of the spectral lines, which are also affected by active regions, as well as spectral activity indicators which trace chromospheric emission in the cores of strong lines. \citet{2015MNRAS.452.2269R} developed a GP framework, schematically illustrated in \autoref{fig:RV_framework}, to model these activity indicators alongside the RVs, and to disentangle them from planetary signals. Each observed time-series is modelled as a linear combination of some latent variable $G(t)$, its time derivative $G'(t)$, white noise, and (in the case of the RVs only) one or more Keplerian signals. $G(t)$ does not have a direct physical interpretation; it loosely corresponds to $F^2$ in the $FF'$ framework of \citet{2012MNRAS.419.3147A}, so that $G'(t) \propto F(t) F'(t)$, but it is modelled as a (typically quasi-periodic) GP, which makes it possible to write down expressions for the covariance between any pair of observations from the any of the time-series included in the analysis. This results in a global covariance matrix of dimensions $(NM,NM)$, where $N$ is the number of observations and $M$ the number of activity indicators included in the analysis (including the RVs themselves). This framework has proved particularly successful for RV confirmation of transiting planets around young stars \citep{2019MNRAS.490..698B,2022MNRAS.512.3060Z}, enabling the detection of planetary signals almost 50 times smaller than the activity signals.  Open-source implementations of this framework have been published recently \citep{2022MNRAS.509..866B,2022A&A...659A.182D}, and extensions thereof are likely to play a significant role in the analysis of next regeneration RV surveys targeting Earth analogues, which need to achieve even better contrast between activity and planetary signals.

\subsubsection{Granulation and asteroseismology}

\dan{Assigned: DFM}

\subsection{AGN variability}

\dan{Assigned: DFM}

\suz{This section is assigned to DFM but as I started writing something about it for the introduction then decided to move it here I have left it in -- feel free to use this as a starting point or ignore it entirely!.}

The Active Galactic Nuclei (AGN) at the centre of many distant galaxies produce stochastic variability in their optical and radio emission, on timescales ranging from hours to years.

As discussed in Section~\ref{sec:intro}, this is --to the best of our knowledge-- the first area in which GPR makes an appearance in the refereed astronomical literature.

\suz{Refer to the first simulated example in \autoref{sec:quasar} here. Some of the text below may need to move to that Section.}

When the quasar is lensed by an intervening galaxy or galaxy cluster, multiple images can be formed, whose brightness can be monitored individually. Measurements of the time-delays between the resulting light curves can be used to constrain the Hubble constant, $H_0$ \citep{doi:10.1146/annurev.aa.30.090192.001523}. Early measurements of these time-delays in the decade following the discovery of the first multiply imaged quasar, the ``double quasar'' 0957$+$561, were hampered by the lack of a generative model for the light curves. Heuristic methods based on interpolation and cross-correlation were developed, but uncertainties were difficult to derive and results obtained with different instruments and different bandpasses weren't always consistent. \citet{prh92a} derived a simple $\chi^2$ metric under the assumption that the observed light curves are shifted, noisy versions of a single sample a Gaussian process with known covariance. Having proposed an empirical procedure to estimate the covariance function from the data, they then optimize the $\chi^2$ with respect to the time-delay between images, and apply standard methods used to derive an uncertainty. While their method relied on an adhoc method to estimate the covariance matrix, and did not propagate the corresponding errors into the final result, it did resolve the prior discrepancy between the estimates of the time-delay from optical and radio dataets \citet{prh92b}. Even then, the requirement to invert the covariance matrix was a barrier to the wider application of this method. This was addressed a few years later by \citet{pr95}, who introduced introduced a fast algorithm, to invert the covariance matrix for exponential kernels. Updated versions of this algorithm now form the basis for fast GP solvers discussed in Section~\ref{sec:fast}.

Quasar power spectrum estimation references: \citet{2014ApJ...788...33K, 2017MNRAS.470.3027K, 2019PASP..131f3001M, 2022ApJ...936..132Y}

\subsection{Compact objects}
\suz{I thought this new heading would be a neat way to introduce applications to pulsars, non-quasar QPOs and GWs?}

% \subsubsection{Pulsars}

To the best of our knowledge, GPs were first applied to pulsar observations by \citet{2014PhRvD..90j4012V} in the context of pulsar timing studies aimed at detecting gravitational waves. They use a GP to model the noise on series of pulse timing measurements, which are used to search for spacetime distortions caused by the gravitational waves.

More recently, \citet{2016MNRAS.456.1374B,2019MNRAS.488.5702B,2022MNRAS.513.5861S} used GPs to model decades-long radio pulsar observations. In this case the GP is used to model the pulses themselves, and to track long-term changes in pulse periodicity and shape. This is done by evaluating (analytically) the first and second derivatives of the fitted GP and using those to find extrema and inflexion points of the pulse profile, along with the associated uncertainties.

% \subsubsection{Quasi-periodic oscillations}

Quasi-periodic oscillations: \suz{Came across this reference while looking for pulsar work: \citet{2022ApJ...936...17H}}

% \subsubsection{Gravitational waves}
Gravitational waves:
\begin{enumerate}
  \item In GW, they always work in the Fourier domain. Given infinite / periodic data, and stationary noise, this diagonalizes the covariance matrix \citep{2002PhRvD..65l2002A, 2017LRR....20....2R}. The periodicity is achieved by windowing the data then FFTing, but this breaks stationarity. This isn't a sensible assumption for the other examples we've discussed, and it gets tricky here for things like ringdown analysis \citep{2021arXiv210705609I}.
  \item Similarly they're always operating in the uniformly-sampled limit (Toeplitz matrices) \citep{2002PhRvD..65l2002A}.
  \item But the PSD of the noise is very complicated and well captured in the F domain. It would be non-trivial to implement it in the time domain and perhaps computationally intractable.
  \item Detailed discussion of approximating noise as diagonal in Fourier domain when all usual assumptions are not met \citep{UNSER1984231}.
\end{enumerate}

% \subsubsection{Transients}
Transients:
\begin{enumerate}
  \item Using GP interpolation: \citep{2016ApJS..225...31L, 2019AJ....158..257B, 2020ApJ...905...94V, 2021ApJS..255...24V}
\end{enumerate}

\dan{Point that this is the use of GPs for interpolation.}

\section{CHALLENGES, PITFALLS, AND SOLUTIONS}
\label{sec:challenges}

So far, we have primarily discussed the benefits of GP models for astrophysics, but in this section we aim to highlight some challenges faced when using GPs in practice.
The most significant challenge that has so far limited broader adoption GPs for astrophysics is their computational cost and scaling, but as we will see below, several techniques have been developed recently that overcome this limitation, in particular for 1--D datasets that are common in time domain astronomy.
We also briefly discuss some contexts where GP models may overfit, or be otherwise misspecified for specific problems.

\subsection{Scalable GP inference}
\label{sec:fast}

The core computations required for GPR are the likelihood function (\autoref{eq:gp_like}), and predictive distributions (\autoref{eq:pred}).
All of these computations require the solution to a linear system, whose dimension scales with the number of data points $N$.
While this can be solved using optimised linear algebra libraries, possibly including acceleration with a GPU or other highly parallel hardware, the raw computational cost of evaluating a GP model typically scales as the cube of the number of data points $N^3$.
This makes it intractable to use GPs for large datasets, practically restricting their use to datasets with less than a few hundred data points.
General approaches have been developed for scaling GP models to larger datasets, and there are some methods that have seen significant impact within the astronomical literature.

Scalable methods for GPR typically fall into two broad---but sometimes conceptually overlapping---categories: \emph{(1)} approximate, or \emph{(2)} exact on a restricted function space.
Approximate methods typically rely on algorithms from computational linear algebra to construct sparse, low-rank, or otherwise structured approximations to the covariance matrix that can be used to approximately solve the required systems to within a target numerical accuracy.
On the other hand, there exist scalable exact methods that are restricted to a particular class of covariance functions, or data with a specific structure (for example, one-dimensional evenly sampled data).
Approximate methods \citep[e.g.,][]{inducing, george, kissgp, 2015arXiv151101870W} have been widely used in the astrophysics literature thanks, in part, to the existence of publicly-available software (e.g.\ \project{GPy}\footnote{\url{https://github.com/SheffieldML/GPy}}; \project{george}, \citealt{george}; \project{GPyTorch}, \citealt{gpytorch}).
However, within time-domain astronomy, scalable exact methods have been more widely used, since time series data often has more amenable structure.

One simple example of a scalable exact method is applicable to one-dimensional evenly sampled data, where the covariance matrix has ``Toeplitz'' structure \citep[e.g.,][]{toeplitz}, which permits more efficient computations using a fast Fourier transform.
The use of Toeplitz structure for scalable Gaussian noise modeling has primarily been applied for gravitational wave data analysis \citep[e.g.,][]{2020PhRvR...2d3298T, 2021arXiv210705609I}, however, most datasets in astrophysics are not evenly sampled, limiting the applicability of this method.

The most widely-used algorithms for scalable exact GPR in time domain astrophysics are based---often not explicitly---on stochastic differential equations \citep[SDEs;][]{Sarkka:2019}, and typically require that the data be ``sortable''.
As an early example in the astrophysics literature, \citet{pr95} presented an algorithm for GPR with linear scaling with the size of the dataset, when using an exponential kernel with unevenly sampled time series data.
This method was later generalized to a wider range of covariance functions \citep{ambikasaran2015generalized, celerite, 2020AJ....160..240G, 2020A&A...638A..95D, 2022A&A...659A.182D}.
The \project{celerite}\footnote{\url{https://github.com/dfm/celerite}} and \project{S+LEAF}\footnote{\url{https://obswww.unige.ch/~delisle/spleaf/doc}} implementations of these methods have been widely-used for a wide range of applications, including: measuring stellar rotation periods \citep[e.g.,][]{2020MNRAS.492.1008G, 2022AJ....164..115N}; characterizing AGN light curves \citep[e.g.,][]{2022ApJ...936..132Y}; and modeling the light curves of transiting planets \citep[e.g.,][]{2019AJ....158...32K} and eclipsing binaries \citep[e.g.,][]{2020ApJS..250...34C}.
A qualitatively different, and independently developed approach to implementing these models uses a state-space representation for the underlying SDE to enable scalable inference \citep{2014ApJ...788...33K,  2021RNAAS...5..107J, 2022arXiv220709327M}.

\dan{Do we want to add a figure here demonstrating the importance of scaling?} \suz{yes why not if time permits...}

\subsection{Overfitting}

In our experience, a common concern that astrophysicists have when working with GPs is that these flexible models may overfit the observations, and decrease the significance of their results, especially when a GP is used as an effective model for noise sources.
While this is not an unjustified concern, we find that the risk is generally overestimated.
As discussed in \autoref{sec:gp-lsq-link}, the determinant in \autoref{eq:gp_like} penalises GP models with ``unspecific'' hyperparameters, instead preferring to explain as much of the data as possible with the mean model.
In other (but equally informal) terms, this means that a GP's modeling capacity will only be used to describe residuals that are not captured within the mean model's support.

All this being said, there do exist cases where the use of a GP can lead to overfitting. One such example is RV planet searches when the ephemeris of the planet(s) is unknown (i.e.\ excluding transiting planets). Unless the time-sampling of the data is truly excellent, the hyper-parameters of the GP activity model are poorly constrained, and the highly multi-modal nature of the the posterior surface means that standard Bayesian inference and model comparison techniques can be misleading. One example of such a situation is the short-period, Earth-mass planet reported by \citet{2012Natur.491..207D} around $\alpha$\,Centauri\,B, which was shown by \citet{2016MNRAS.456L...6R} to be an artefact of the time-sampling: a small peak in the window function of the observations was serendipitously ``boosted'' by the complex model necessary to account for the star's activity signal. This is a concern with any flexible model applied to a sparse dataset, but it can be less obvious for GPs, which can be very flexible with comparatively few hyper-parameters. Useful strategies to deal with this kind of problem include cross-validation (as discussed in Section~\ref{sec:model_selection}) and performing exhaustive injection-recovery tests to build (or undermine, as the case may be!) confidence in one's results.

\subsection{Model misspecification}

\dan{Outliers. Heavy tails.}

% Including when not to use GPs.

% In this section we will discuss some of the main drawbacks of using Gaussian processes namely computational cost and overfitting. We will discuss how recent advances are helping overcome the former, and illustrate how the latter arises using examples from the astronomical literature. We will suggest methods for diagnosing and avoiding overfitting (e.g. cross validation). Estimated length: 5 pages

\section{OPEN-SOURCE GAUSSIAN PROCESS SOFTWARE}
\label{sec:open}

It is reasonably straightforward to implement a simple GP model in code, and within astrophysics it has been common for authors to implement custom GPs for their analysis.
However, things get significantly more complicated when implementing the scalable or approximate methods discussed in the previous section.
Similarly, it can be tedious to experiment with different kernel functions and inference methods without building a non-trivial modeling infrastructure, something that has typically been ad-hoc in astrophysics research.

Luckily, driven by the immense popularity of GPR in machine learning, the physical sciences, and other fields, there are a plethora of open source tools that have been developed to simplify this process for a wide range of applications.
Many of these tools are designed for scalability, flexibility, and ease of use.
In this section, we describe some popular libraries in this space, while cautioning the reader that this is not a comprehensive list, and this domain changes quickly, so the discussion may become outdated more quickly than the rest of this document.
Most of our discussion will focus on tools implemented in the \project{Python} programming language since it is---at the time of writing---the most popular language, both in astronomy and GP modeling, but there are tools available in all other popular languages.

Alongside this review, we have released all the source code to generate the figures in this document using the \project{tinygp} library\footnote{\url{https://tinygp.readthedocs.io}} \citep{tinygp}, which is a relatively new open source \project{Python} library for GP modeling.
While this library hasn't yet been widely used in the astrophysics literature, we believe\footnote{Note that the development of the \project{tinyp} library is led by D.\ Foreman-Mackey, an author of this review.} that it is a good choice for many applications.
\project{tinygp} uses the \project{JAX} library as its computational backend, giving it good computational performance (including support for GPU hardware) with a pure-\project{Python} user interface.
\project{JAX} also provides support for automatic differentiation \citep[e.g.,][]{autodiff}, which can be used to accelerate all the inference schemes described in Sections~\ref{sec:inference} and~\ref{sec:model_selection}.

Perhaps the most widely used \project{Python} library for GP inference in astrophysics is \project{george}\footnote{\url{https://george.readthedocs.io}} \citep{george}, which is a relatively simple library that implements most of the standard kernel functions described in \autoref{sec:kernels}, and an approximate solver that can be used to scale to moderately large ($N \sim 1000$) datasets.
More recently, the scalable methods implemented by \project{celerite}\footnote{\url{https://github.com/dfm/celerite}} \citep{celerite} and \project{S+LEAF}\footnote{\url{https://obswww.unige.ch/~delisle/spleaf/doc}} \citep{2020A&A...638A..95D, 2022A&A...659A.182D} have become popular for time domain GP models in astrophysics.
These libraries require one-dimensional---or appropriately structured \citep{2020AJ....160..240G, 2022A&A...659A.182D}---input data, and a specific set of kernel functions, although these kernels form a complete \citep{JMLR:v22:21-0072} and physically interpretable \citep{celerite} basis, and it has recently been demonstrated that the Mat\'ern class of kernels can be exactly represented \citep{2021RNAAS...5..107J}.

Outside of astrophysics, the \project{GPy}\footnote{\url{https://github.com/SheffieldML/GPy}} and \project{GPyTorch}\footnote{\url{https://gpytorch.ai}} \citep{gpytorch} projects provide mature and widely used GP implementations.
\project{GPyTorch}, in particular, is designed for scaling GP modeling to large datasets, using approximate methods \citep[e.g.,][]{kissgp, 2015arXiv151101870W} that do not have the same restrictions as \project{celerite} and \project{S+LEAF}.
\project{GPyTorch} has been used for a variety of applications in astrophysics \citep[e.g.,][]{2022A&A...658A.166D, 2022MNRAS.511.5597L}, but not yet as widely as the other libraries discussed so far, perhaps because it does not integrate into the astronomical workflow as easily as the projects that were designed with astrophysicists in mind.

\section{CONCLUSIONS}
\label{sec:concl}

Gaussian Process Regression has become a key element in the toolkit of modern astronomers, particularly those working with time-domain datasets. It offers a principled approach to modelling stochastic signals and nuisances, particularly correlated noise, whose importance has grown significantly in the past decade with the development of major precision time-domain monitoring surveys (e.g.\ planetary transit surveys). GPR is explicitly 
rooted in probability theory, making it readily interpretable and enabling astronomers to quantify their confidence (or lack thereof) in the results of their analysis. Despite the simplicity of the underlying mathematical framework, even basic GP models are extremely flexible, and a small handful of commonly used ``building blocks" have been used successfully to explain the observed variability of astronomical objects on all scales. On the other hand, this flexibility comes at a price: a ``good fit'' with a particular model is no longer a fail-safe indication that this particular model is ``the correct one'', so users must proceed with extra caution.

\subsection{Summary of the review}

After a very brief historical introduction, we started this review by introducing two simulated examples, which demonstrate the two main classes of use for GPR in astronomy, namely to model a nuisance signal (in the case of a planetary transit) or a stochastic signal of interest (in the quasar time-delay example). The source code for these and other worked examples in the review is available alongside this review in the form of {\sc Jupyter} notebooks \suz{[give details]}, and interested users are encouraged to uses these to further their understanding of the methodology and potentially adapt them to their own datasets. Next, we gave a succinct introduction to the key equations underlying GPR regression, and sketched out a typical workflow for applying GPR to a new dataset. Like any data analysis method, GPR requires the user to make a number of choices, the most important of which is the choice of covariance function. We listed some of the most widely used covariance functions, outlined how to construct more complex ones, and gave practical advice on how to assess the validity of one's modelling choices. These strategies include standard Bayesian model comparison but also cross-validation, which can be more appropriate in cases where predictive power is important.

We then reviewed the growing literature on applications of GPR to time-domain datasets, from exoplanet detection and characterisation, where GPs are mostly used to account for instrumental or stellar nuisance signals, to the characterisation of stellar variability itself, where the GP is now used to model a stochastic signal for its own sake. These examples are drawn from the exoplanet and stellar astrophysics literature, which we are most familiar with, but we also attempted to give an overview of recent applications of GPR to two other classes of astronomical objects where it is increasingly important, namely AGN and compact objects. Altogether, these applications include a very diverse array of GP models with varying levels of complexity, including multi-dimensional inputs and outputs, derivative and integral observations, and applications where the GP model is physically motivated versus cases where it is merely an effective model. We noted how GPR can be used to tease apart the deterministic and stochastic components of a model, but also different components of a stochastic process.

Finally, we discussed the challenges and pitfalls of GPR, noting recent algorithmic and technical developments that have brought down their computational cost, once prohibitive for large datasets, and outlining some strategies to limit the risk of overfitting. Finally we reviewed the ecology of free software resources for GPR, which make it easy to implement a t test a GP model for a new problem in just a few lines of code.

\subsection{Future perspectives}

In certain application domains, such as the analysis of photometric and radial velocity time-series from exoplanet surveys, GPR is already used fairly widely. Its usefulness as an effective model for astrophysical or instrumental nuisance signals is well established, and it is seen as a fairly standard element of the astronomer's data analysis toolkit. In these existing applications, we foresee three main routes for further progress in the near future. One is the development of more physically motivated (as opposed to merely effective) GP models for stellar variability \citep[see e.g.][]{2021AJ....162..124L} \suz{[add asteroseismology refs]}, so that GPR can be used to characterise and understand these stellar signals for their own sake. The other is a more systematic use of GPR as part of the detrending and detection pipelines for exoplanet transit and RV surveys. Computational cost considerations would have made this impractical just a few years ago, but this obstacle has been lifted thanks to the recent development of scalable GP models. Finally, as JWST begins to deliver much exoplanet atmosphere observations with much higher signal-to-noise (including, importantly, more precise and relevant housekeeping parameters), a new generation of systematics models will no doubt be developed to get the most out of these observations and it seems likely that GPR would have a role to play in that.

In other applications, such as pulsars, QPOs and astrophysical transients, GPR remains more of a niche pursuit so far, but promising early results should motivate further work in the next few years. The next decade will see a vast increase in the scale of the relevant datasets, with major surveys such as the Vera Rubin Observatory's Legacy Survey of Space and Time (LSST) or the Square Kilometer Array (SKA) routinely delivering tens of TB per night. Even with the recent development of scalable GP models, it is unclear whether GPR cam be useful as a ``first contact'' analysis tool for these very large surveys, but it has clear potential for the analysis of specific subsets of data.

One area of active development concerns the problem of non-Gaussianity. While the assumption of Gaussianity implicit to GPR is often good enough in practice, it is almost never strictly correct, and this can become problematic when extreme precision is required on the model hyper-parameters, or when the data contain (a relatively small proportion of) significant outliers. Stochastic process models based on distributions with higher kurtosis than the Gaussian, for example Student's-$t$ processes \citep[see e.g.][]{Tracey_2018} are an interesting alternative, though they have not been extensively used in astronomy to our knowledge.

Another area where GPR could become more widely used in future is for intelligent observation planning: if a set of variable objects is being monitored as part of a survey, using a GP to model existing observations would enable the observer to make predictions (including uncertainties) for future observations, and therefore to prioritise among the target list in real time according to some metric of choice. \suz{Keep or remove? I'm not sure we have anything truly meaningful to say here.}

\suz{It would be nice to include a few words about new kinds of GP models bridging the gap between "traditional models" and AI, like Gaussian Process networks [ref recent RV paper] - but I am not sure I know enough about them.}

% \section{FIRST-LEVEL HEADING}
% This is dummy text.
% % Heading 2
% \subsection{Second-Level Heading}
% This is dummy text. This is dummy text. This is dummy text. This is dummy text.

% % Heading 3
% \subsubsection{Third-Level Heading}
% This is dummy text. This is dummy text. This is dummy text. This is dummy text.

% % Heading 4
% \paragraph{Fourth-Level Heading} Fourth-level headings are placed as part of the paragraph.

% %Example of a Figure
% \section{ELEMENTS\ OF\ THE\ MANUSCRIPT}
% \subsection{Figures}Figures should be cited in the main text in chronological order. This is dummy text with a citation to the first figure (\textbf{Figure \ref{fig1}}). Citations to \textbf{Figure \ref{fig1}} (and other figures) will be bold.

% \begin{figure}[ht]
%   \includegraphics[width=3in]{SampleFigure}
%   \caption{Figure caption with descriptions of parts a and b}
%   \label{fig1}
% \end{figure}

% % Example of a Table
% \subsection{Tables} Tables should also be cited in the main text in chronological order (\textbf {Table \ref{tab1}}).

% \begin{table}[ht]
%   \tabcolsep7.5pt
%   \caption{Table caption}
%   \label{tab1}
%   \begin{center}
%     \begin{tabular}{@{}l|c|c|c|c@{}}
%       \hline
%       Head 1              &          &                   &         & Head 5    \\
%       {(}units)$^{\rm a}$ & Head 2   & Head 3            & Head 4  & {(}units) \\
%       \hline
%       Column 1            & Column 2 & Column3$^{\rm b}$ & Column4 & Column    \\
%       Column 1            & Column 2 & Column3           & Column4 & Column    \\
%       Column 1            & Column 2 & Column3           & Column4 & Column    \\
%       Column 1            & Column 2 & Column3           & Column4 & Column    \\
%       \hline
%     \end{tabular}
%   \end{center}
%   \begin{tabnote}
%     $^{\rm a}$Table footnote; $^{\rm b}$second table footnote.
%   \end{tabnote}
% \end{table}

% % Example of lists
% \subsection{Lists and Extracts} Here is an example of a numbered list:
% \begin{enumerate}
%   \item List entry number 1,
%   \item List entry number 2,
%   \item List entry number 3,\item List entry number 4, and
%   \item List entry number 5.
% \end{enumerate}

% Here is an example of a extract.
% \begin{extract}
%   This is an example text of quote or extract.
%   This is an example text of quote or extract.
% \end{extract}

% \subsection{Sidebars and Margin Notes}
% % Margin Note
% \begin{armarginnote}[]
%   \entry{Term A}{definition}
%   \entry{Term B}{definition}
%   \entry{Term C}{defintion}
% \end{armarginnote}

% \begin{textbox}[ht]\section{SIDEBARS}
%   Sidebar text goes here.
%   \subsection{Sidebar Second-Level Heading}
%   More text goes here.\subsubsection{Sidebar third-level heading}
%   Text goes here.\end{textbox}

% \subsection{Equations}
% % Example of a single-line equation
% \begin{equation}
%   a = b \ {\rm ((Single\ Equation\ Numbered))}
% \end{equation}
% %Example of multiple-line equation
% Equations can also be multiple lines as shown in Equations 2 and 3.
% \begin{eqnarray}
%   c = 0 \ {\rm ((Multiple\  Lines, \ Numbered))}\\
%   ac = 0 \ {\rm ((Multiple \ Lines, \ Numbered))}
% \end{eqnarray}

% Summary Points
\begin{summary}[SUMMARY POINTS]
  \begin{enumerate}
    \item Summary point 1. These should be full sentences.
    \item Summary point 2. These should be full sentences.
    \item Summary point 3. These should be full sentences.
    \item Summary point 4. These should be full sentences.
  \end{enumerate}
\end{summary}

% Future Issues
\begin{issues}[FUTURE ISSUES]
  \begin{enumerate}
    \item Future issue 1. These should be full sentences.
    \item Future issue 2. These should be full sentences.
    \item Future issue 3. These should be full sentences.
    \item Future issue 4. These should be full sentences.
  \end{enumerate}
\end{issues}

%Disclosure
\section*{DISCLOSURE STATEMENT}
The authors are not aware of any affiliations, memberships, funding, or financial holdings that
might be perceived as affecting the objectivity of this review.

% Acknowledgements
\section*{ACKNOWLEDGMENTS}
S.\ A.\ wishes to acknowledge support from the European Research Council (ERC) under the 
European Union's Horizon 2020 
research and innovation programme (Grant agreement No. 865624).

\dan{Add Dan's acknowledgements}

% References
%
% Margin notes within bibliography
%\section*{LITERATURE\ CITED}

\bibliographystyle{ar-style2}
\bibliography{bib}

\section*{RELATED RESOURCES}

SA gave a set of lectures on GPR for time-domain astronomy as part of the 2021 \emph{Saas Fee Advanced Course in Astrophysics}, for which PDF lecture notes are available at \url{shorturl.at/hrstz}, along with {\sc Python} exercises in the form of {\sc Jupyter} notebooks. \suz{This simple URL reference will need to be replaced with a citation to the official lecture notes published by Springer if those appear before the present review.}

\end{document}